### 2012-12-02 CMPT771/project_report/paper.tex
initial commit

%File: formatting-instruction.tex
\documentclass[letterpaper]{article}
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\frenchspacing
\pdfinfo{
/Title (Formatting Instructions for Authors Using LaTeX)
/Subject (AAAI Publications)
/Author (AAAI Press)}
\setcounter{secnumdepth}{1}  
 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{Probability Based Partitioning of Online Social Networks}
\author{Vivek Mahajan\\
vmahajan@sfu.ca\\
}
\maketitle
\begin{abstract}
\begin{quote}
AAAI creates proceedings, working notes, and technical reports directly from electronic source furnished by the authors. To ensure that all papers in the publication have a uniform appearance, authors must adhere to the following instructions. 
\end{quote}
\end{abstract}

\section{Introduction}

\section{Related Works}

\section{Proposed Idea}

\cite{test123}
\bibliographystyle{aaai} \bibliography{testbib}
\end{document}

### 2012-12-02 CMPT771/survey_report/aaai_script.sh

@@ -1,0 +1,115 @@
+#!/bin/bash
+
+# ----------------------------------------------------------------------
+# This shell script produces the latex source-package of a paper
+# as required by AAAI, in preparation for printed proceedings.
+# Copyright (C) 2009 Christian Fritz "fritz at cs dot toronto dot
+# edu"
+# ----------------------------------------------------------------------
+# This program is free software: you can redistribute it and/or
+# modify it under the terms of the GNU General Public License as
+# published by the Free Software Foundation, either version 3 of
+# the License, or (at your option) any later version. 
+#
+# This program is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE. See the
+# GNU General Public License for more details.
+#
+# You should have received a copy of the GNU General Public
+# License along with this program. If not, see
+# <http://www.gnu.org/licenses/>.
+# ----------------------------------------------------------------------
+
+# function for recursive processing of style files
+recurseStyles() {
+    echo "..recursively processing required packages of $1"
+    PACKAGES=`grep RequirePackage $1 | sed 's/RequirePackage\[[^]]*\]/RequirePackage/g' | sed 's/.*RequirePackage{\([^}]*\)}.*/\1/' | sed -e 's/, */\n/g'`
+    for name in $PACKAGES; do
+	if [ ! -e sources/$name.sty ]; then 
+	    echo "..locating and copying: $name"
+	    FILE=`locate /$name.sty | head -n 1`
+	    if [ -e $FILE ]; then
+		cp $FILE sources
+	    else
+		echo "..cannot locate $name.sty"
+	    fi
+	    if [ -e sources/$name.sty ]; then
+		recurseStyles sources/$name.sty
+	    fi
+	fi
+    done
+}
+
+#            -------------------------------              
+
+# give a tex file as parameter
+if (( $# < 1 )); then
[…]

### 2012-12-12 CMPT771/survey_report/paper.tex
initial commit
adding abstract and objective function section to the survey report

--- 2012-12-02 CMPT771/project_report/paper.tex
@@
 \frenchspacing
-\pdfinfo{
-/Title (Formatting Instructions for Authors Using LaTeX)
-/Subject (AAAI Publications)
-/Author (AAAI Press)}
+%\pdfinfo{
+%/Title (Formatting Instructions for Authors Using LaTeX)
+%/Subject (AAAI Publications)
+%/Author (AAAI Press)}
 \setcounter{secnumdepth}{1}  
@@
 %
-\title{Probability Based Partitioning of Online Social Networks}
+\title{Finding communities in graphs}
 \author{Vivek Mahajan\\
@@
 \begin{quote}
-AAAI creates proceedings, working notes, and technical reports directly from electronic source furnished by the authors.
-To ensure that all papers in the publication have a uniform appearance, authors must adhere to the following instructions.
+Detecting clusters or communities in large real-world graphs such as large social or information networks is a problem of considerable interest. In general, it is the 
+organization of vertices in clusters, with many edges joining vertices of the same cluster and comparatively few edges joining vertices of different clusters. 
+In practice, one typically chooses an objective function that captures the intution as given by the definition above, and then one applies approximation algorithms
+or heuristics to extract sets of nodes which appears to be "good" with respect to that objective function.
+
+In this survey, we will start with the applications of clustering in various domains. We will describle various objective functions and will also discuss some of the
+most cited community finding papers.
+
 \end{quote}
@@
 
-\section{Related Works}
+\section{Objective Functions}
+If you have clusters of a given graph as an output of a clustering algorithm, we need to have some quantitative measure of how "good" they are. We call them objective
+function which the algorithms try to optimize. The intution behind most of the objective functions is that a good cluster has many edges internally and few edges 
+pointing outside. In this section, we look at the other objective functions that are based on this intution.
 
-\section{Proposed Idea}
+In general, the two criterias contributes heavily to the formulation of objective function. The first is the number of edges between the members of the clusters, and 
+second is the number of edges between the members of the cluster and remainder of the network. So, based on whether an objective function is using either one or both 
+criteria in its formulation, we can divide the objective functions into two categories\cite{leskovec2010}. The first group, that we refer to as Multi-criterion scores, combines both 
+criteria into a single objective function; while the second group of objective functions uses only  a single of the two criteria. We will give a quick overview of the
+various definitions of the objective functions which falls under the two categories.
+
+\subsection{Multi-criterion scores}
+Let $G(V,E)$ be an undirected graph with $n=|V|$ and $m=|E|$ edges. Let S be the set of nodes in the cluster, where $n_S$ is the number of nodes in S,
+$n_S=|S|$; $m_S$ the number of edges in $S$,  $m_S=|\{(u,v):u\in S, v\notin S\}|;$ and $d(u)$ is the degree of node $u$.
+
+We consider the following metrics $f(S)$ that capture the notion of a quality of the cluster. Lower value of score $f(S)$ signifies a more community-like set of 
+nodes.
+\begin{itemize}
+\item \textbf{Expansion}:  
+\item \textbf{Internal density}:  
+\item \textbf{Cut Ratio}:  
+\item \textbf{Normalized Cut}:  
+\item \textbf{Maximum-ODF(Out Degree Fraction)}  
+\item \textbf{Average-ODF}:  
+\item \textbf{Flake-ODF}: 
+\end{itemize}
+
+\subsection{Single-criterion scores}
+\begin{itemize}
+\item \textbf{Modularity}:
+\item \textbf{Modularity ratio}:
+\item \textbf{Volume}:
+\item \textbf{Edges cut}:
+\end{itemize}
+
+\section{Algorithms}
 
 \cite{test123}
-\bibliographystyle{aaai} \bibliography{testbib}
+\bibliographystyle{aaai} \bibliography{bibfile}
 \end{document}

### 2012-12-13 CMPT771/project_report/paper.tex
updating project report

@@
 \begin{quote}
-AAAI creates proceedings, working notes, and technical reports directly from electronic source furnished by the authors.
-To ensure that all papers in the publication have a uniform appearance, authors must adhere to the following instructions.
+Online social networking has  become a phenomenon in the last
+few years. The growing usage of social networks has put the service providers
+under a lot of pressure to efficiently process the users queries.
+ In social networks like orkut, most of the queries access the immediate neighbors of the current node and
+neighbors of its neighbors.  Existing graph partitioning approaches focus on reducing the number
+ of inter server links, neglecting the query
+patterns. Our algorithm focuses on partitioning the graph so as to
+minimize the number of servers to be accessed in order to
+ process a query. Our results show that the proposed algorithm gives useful partitioning of social networks,
+thereby reducing inter server accesses.
+
 \end{quote}
@@
 \section{Introduction}
+    In the last few years, the online social networks(OSNs) have grown in huge proportion with respect to the number of users as well as its utility.
+The amount of usage of OSNs exceeds e-mail usage nowadays, which is remarkable.
+ OSNs are the most popular sites on the Internet.
+A study\footnote{http://www.news.com.au/technology/story/0,28348,25170441-5014239,00.html} reported that OSNs account for 10\% of all the Internet time worldwide.
+The growth can be appreciated by the fact that there was 1382\% increase in the user base of twitter in one month[Feb-Mar 09] \footnote{http://blog.nielsen.com/nielsenwire/onlinemobile/twitters-tweet-smell-of-success}.
+There can be variations of OSNs in terms of usability e.g On orkut, the users interact with each other by exchanging scraps or photographs, while on SLASHDOT, the users interact via blogs.
+Facebook, LinkedIn are other examples of OSNs.
+
+    The phenomenal growth in the size of OSNs has given rise to scalability issues that can be solved by distributing online social network graph among different servers.
+However the basis of partitioning  is a major issue.
+That is, the graph partitioning could be based on structural, geo-political or traffic properties, which are discussed in detail in \cite{arxviv} for twitter and orkut.
+In this paper, we introduce a new partitioning criterion called \texttt{hop based partitioning} which takes into consideration the typical nature of social network queries.
+
+\subsection{Hop Based Partitioning}
+
+Social network queries typically visit their one hop and two hop neighbors.
+The small world phenomenon links any two users on a social network with a distance of six hops.
+Hence, while partitioning a social network, it would be beneficial to maintain the one-hop and two-hop neighbors of a frequented query node on the same server\footnote{through the rest of the paper the server and partition terms are used interchangeably} as itself.
+Further, it is expected that any graph partitioning algorithm attempts to reduce the inter server edges and equally distribute the data across servers.
+
+For example, \cite{flickr} analyzed large-scale traces of information
+dissemination in the Flickr social network. Flickr allows users to create two types of links: links to favorite
+photos (called favorites in Flickr) and links to other users (called contacts in Flickr). Users may \texttt{favorite-mark} a photo
+ and share interesting photos with others.
+\cite{flickr} refers to users who include a photo in their favorite photos list as fans of that photo.
+It was observed that for less popular photos, 91\% of all fans are within two hops of the uploaders.
+Further, for even for top popular photos, 81\% of all fans are within two hops of uploaders.
+\cite{OHM} mentions that most information is found one-hop away from the query node.
+According to a study\cite{4}, the maximum distance between two persons is six hops.
+Hence, it is a reasonable assumption that a user interacts primarily within his 2-3 hop radius and at most his six hop radius.
+
+    In this project, we propose a novel probability based objective function and  an algorithm to find 
+the clusters which optimizes that. Given a social network graph $G$, our algorithm partitions the graph while 
+meeting the following constraints:
+\begin{itemize}
+\item For any node in a partition most of its two-hop neighbors lie within the same partition.
+\item All partitions are approximately equally sized.
+\end{itemize}
+
+The paper is organized as follows. The related
+works is shown in section 2, the objective function in exaplained in section 3, We describe our algorithm  in
+ section 4 and  experimental results are shown in section 5.
+ Finally, we conclude the paper in section 6.
 
 \section{Related Works}
+  The study of graph partitioning algorithms based on different properties of social networks has been proposed by \cite{arxviv}.
+In the recent past, graph partitioning for dealing with scalability issues and handling large datasets has been proposed by \cite{5,6}.
 
-\section{Proposed Idea}
+    With the increased usage of social networks, the scientists have become more interested in analyzing the properties of social networks. 
+Sociologists showed that the average path distance between two Americans is 6 hops \cite{4} and \cite{8} shows
+the small world effect analysis. 
+Study of a online social network  finds the small world behavior and local clustering
+ property  \cite{9}.
+ The small world property for offline social networks has been discussed in \cite{9,10,11} 
+and it was verified for the online social network \cite{12,13}. Analysis of different structural and scale free properties has also been extensively analyzed \cite{12}.
 
-\cite{test123}
-\bibliographystyle{aaai} \bibliography{testbib}
+    The traditional graph partitioning algorithms\cite{16,17,18} partition the graph into roughly equal sized partition  and focusses on minimizing the edge cuts.
+ Further, the sociological approaches like hierarchical clustering based partitioning and natural division 
+based partitioning approaches are popularity known\cite{19}. Another genre of graph partitioning 
+algorithms that focus on producing quality partitions in less time are multilevel algorithms, which 
+include METIS\cite{20} and its modification for power law graphs\cite{21}. There are other partitioning 
+algorithms like Kernighan-Lin\cite{22} and it's linear time heuristics\cite{23} can also be used as 
+refinement algorithms.
+
+    In this paper, we  compare our results with METIS\cite{20}.
+METIS focuses on minimizing the cuts between the partitions as well as into roughly equal partitions.
+The METIS algorithm is divided into three phases termed as coarsening, partitioning and uncoarsening/refinement.
+ The main motivation behind the algorithm is to generate an approximate graph of the original graph, partitioning it and then retracting the coarsening phase while doing the refinement between each iteration.
+ Our algorithm also follows a hierarchical clustering approach where repeated merging of clusters is done
+ for some iteration and then the graph randomly partitioned with imposed contraints of having nearly equal 
+partitions. Finally, the refinement algorithm
+ is applied on the clusters thus obtained. We differ from METIS, in terms of our partitioning criterion, 
+our novel way of clustering and partitioning, and  the modification and the method of refinement applied 
+which is applied only once.
+
+\section{Probability Based Objective Function}
+
+   To calculate our objective function, we assume that there is an equal chance of each user of an 
+OSN to come online. However, we can easily modify our objective function if we know the probability of each user
+coming online.
+
+ Given some partitions, an evaluation criteria should be able to give a good index of 
+the quality of partitions in terms of fraction of one and two hops lying
+outside the partition and should also give importance to the past interaction between the users of OSNs.
+
+This index calculates the probability of one and two hops from a node to lie in different partition.
+
+Formally, let $X$ = $\{x_1,x_2,x_3,\cdot\cdot\cdot,x_n\}$ be the nodes in an OSN, $x_i$ = $i^{th}\ node$, 
+$\{y_1, y_2, y_3,\cdot\cdot\cdot,y_j\}$ be the $j$ neighbors of $x_i$. Let $\{v_1, v_2, v_3, \cdot\cdot\cdot,
+ v_j\}$ be the number of interactions between $x_i$ and its neighbors. And $V_i = \sum_{k\in j} v_k$.
+ Let $Y^\prime$ be 
+the subset of neighbors of $x_i$ which lie in partition other than $x_i$'s.
+Let $Y^\prime = \{y_1^\prime, y_2^\prime, \cdot\cdot, y_k^\prime\}$.
+Let the interaction between $x_i$ and the users in this subset be  $V^\prime = \{v_1^\prime, v_2^\prime, \cdot\cdot, v_k^\prime\}$.
+$V_i^\prime = \sum_{p\in k} v_p^\prime$.
+
+        Let $P_{x_i}^{1-hop}$ be the probability of one-hop neighbor of $x_i$ to lie in separate partition, and $P_{x_i}^{2-hop}$ be probability of two hops from $x_i^{th}$ node to lie in the separate partition.\\
+        Clearly, $P_{x_i}^{1-hop}= \frac{V_i^\prime}{V_i}$ \\
+        $P_{x_i}^{2-hop}=\frac{1}{Number\ of\ neighbors\ of\ x_i} * \displaystyle\sum_{k=1}^j P_{x_i y_k}$, \\
+        where,\\
+        $P_{x_i y_k}$ = $\frac{v_k}{V_i}$* $P_{x_{k}}^{1-hop}$ =  Probability of two hop from  $x_{i}$ to lie  in a  partition other than  its own through $y_k$, if $x_i$ and $y_{k}$ are in the same partition.\\
+        $P_{x_i y_k}$ =  1 , if $x_i$ and $y_k$ are in the separate partition, as we assume if the one hop is on a different partition, it is almost
+certain that two hop lie in the same partition.\\
+        $Probability\ Index=\frac{1}{Total\ number\ of\ nodes}*\displaystyle\sum_{i=1}^nP_{x_i}^{2-hop}$, $n$ is the total number of nodes.\\
+
+        This probability index gives us an overall probability of a two hop query from a node to lie in the
+ different partition. In the following section we will propose an algorithm which optimizes this objective 
+function.
+
+\section{Our Algorithm}
+\begin{table*}[htb]
+\centering
+\begin{tabular}{|l|}%putting \begin{tabular}{|l|} gives table with side-lines.
+\hline
+\textbf{Our Algorithm}\\
+\hline
+\textbf{Input: }Graph, No of Partitions, No. of iterations\\
+\hline
+\textbf{Phase 1 - Graph Coarsening }\\
+1. Node weight calculation$()$ - Calculate the weights for each node according \\
+~~~ to the algorithm discussed in section 3.2\\
+2. Clustering$()$ - Forming clusters according to the sorted order\\
+~~~ of the weights.\\
+3. Graph formation from clusters$()$ - Consider a cluster as a node for the\\
+~~~ graph and collapse the edges to form a graph\\
+3. Repeat step 1-3 for the desired number of iterations.\\
+\textbf{Phase 2 - Partitioning}\\
+ Randomly allocate nodes into the desired number of clusters. We can use any traditional\\
+ approach towards graph partitioning, in our algorithm  we are using a randomized one\\
+
+\textbf{Phase 2 - Refining}\\
+ Refine the partitions using kenighan-Lin algorithm or the modified one.\\
+\hline
+\end{tabular}
+%\vskip -0.1in
+\caption{Our Algorithm}
+\label{alg}
+\end{table*}
+
+In this section, we introduce our graph partitioning algorithm.
+A social network  partitioning algorithm is expected to reduce inter server accesses requiring neighbors and the neighbors of neighbors of a query nodes to be present in the same server as itself.
+ To reduce the partitioning overhead, it is important that inter server links be reduced while trying to maximize the number of nodes for which its one and two hop neighbors be on the same server as itself.
+ Further, in order to equally distribute the workload across servers and optimally use the servers used, it is
+ essential that the partitions size should be approximately equal. Our algorithm attempts to fulfil 
+ the above requirements while attempting to maximize the number of nodes for which a majority of its 
+one-hop neighbors and with next priority its 2-hop neighbors are included within the same server.
+ Our algorithm is divided into three phases, namely the graph coarsening phase, partitioning 
+phase and the refinement phase. The graph coarsening phase reduces the size of the original graph by some
+factor by clustering the nodes \emph{intelligently}.  A weight function is used 
+in order to decide which nodes should be clustered  during the coarsening phase. In the partitioning phase, we 
+randomly generate partitions of the coarsened graph. The refinement phase next 
+reduces the inter partition edges by making slight modifications to the partitioned graph. We next give 
+definitions of commonly used terms followed by a detailed description of each of the phases.
+
+\subsection{Definitions}
+\textbf {\it{Social Network Graph :} }The users in the social network
+form the nodes of the graph while the relationships between
+the users form the edges of the social network graph.\\
+\textbf{\it{k-hop neighbor :}} A node is said to be a k-hop neighbor of a
+query node if there are  $K$-edges separating the node
+from the query node. A one hop neighbor can also be a two hop neighbor and so on in case of triangles present in the graph
+
+\subsection{Coarsening Phase}
+
+The aim of this phase is to produce an approximate partitioning of the original graph. Our approach is a heirarchical one. 
+We start with the original graph and then we try to reduce the graph size by \emph{intelligently} clustering
+nodes at each level. After clustering nodes, we form a new graph with clusters from the previous level treated
+as nodes in the current level. After repeating the steps for some iterations, we apply a traditional graph
+partitioning algorithm to \emph{latest} graph. And then we do the refining on the produced clusters to get
+the desired output. In the following sections, we will describe our \emph{intelligent} clustering approach at
+each level and then we will continue our discussion with the explanation of the refining algorithm.
+\begin{itemize} % 1
+
+\item \textbf{Graph Definition:}
+
+Graph  $G$=$(V,E)$, where $V$ = $\{v_1,v_2,v_3,\cdot\cdot\cdot,v_n\}$ ,$v_i$ represents a cluster of nodes. 
+In the first level, $v_i$ consists of only single node and $E$ is the edge set which consists of $e_{ij}$ i.e the number of interactions between the $v_i$ and $v_j$ i.e for any $v_i$ and $v_j$,if there are $k$ edges between
+the two then $e_{ij}$ will be equal to $k$, for the first level $e_{ij}=number\ of\ interactions\ between\ v_i\ and\ v_j$.\footnote{number of interactions might the average number of times user $v_i$ and $v_j$ have interacted
+with eachother}
+
+\end{itemize} % 1
+
+\begin{itemize} % 1
+
+\item \textbf{Node Weight Calculation:} In this project, we adopt a probability based weight function. 
+Nodes of higher weight values are later picked as seeds during the process of clustering.
+
+\begin{itemize} %2
+
+\item \textbf{Probability based weight function:}  
+In section 3, we described the $P_{x_i}^{2-hop}$ as the probability of two hops from $x_i^{th}$ to lie in separate
+partition. To calculate the weight for each node $x_i$, we calculate $P_{x_i}^{2-hop}$ assuming the nodes and its 
+neighbors are in one cluster and rest of the nodes are in separate cluster. Higher the weight, more are the
+chances that even if we have the node and its neighbors in one cluster, still there is high chance of two hops
+from node $x_i$ lying in separate cluster. So we start by clustering $\emph{low weight}$ nodes and we continue
+ in ascending order.
+
+\end{itemize} % 2
+
+\item \textbf{Clustering Algorithm:} The algorithm behaves like a hierarchical clustering algorithm where at each level clusters are formed by clustering the clusters of the previous level.
+The clustering algorithm varies based on the weight picking function used as described below.
+
+For a given level $i$, the nodes of the graph at level $i$ are processed in decreasing order of weight.
+Initially the input graph forms the input of the clustering algorithm for level one.
+For each level $i >1$, a new graph is generated which is fed as input to the clustering algorithm.
+
+\begin{itemize} % 2
+
+\item \textbf{Probability based clustering:} Each selected node assigns all its immediate neighbors to a new cluster. 
+A neighbor that has already been assigned to another cluster in the same iteration of level is ignored. 
+The process is repeated until every node is assigned some cluster number.
+
+\end{itemize} % 2
+
+A predefined cluster size is set as:\\
+
+$Cluster$ $Size$ = ($Number$ $of$ $Nodes$ / $Number$ $of$ $partitions$) $+$/$-$ $deviation$ $value$.\\
+
+where $deviation$ $value$= $x\%$ of ($Number$ $of$ $Nodes$ / $Number$ $of$ $partitions$) for some user defined value of $x$.\\
+
+The user defined value $x$ in the equation above is referred to as the "limit value" through the rest of the paper.
+
+The size of any cluster is required to be less than the predefined $Cluster$ $Size$.
+If no node can be selected as a seed node as on selecting it would be giving rise to a cluster that exceeds the $cluster$ $size$, then some heuristics applied to cluster the remaining nodes in a graph.
+
+A cluster that cannot accommodate further nodes is marked and the new graph formed ignores the node representing this cluster for further clustering.
+
+\item \textbf{Graph Formation from Clusters:} For the first level before the algorithm starts, each node of the input graph is represented as a cluster in itself.
+After the first level of clustering at level one, a new graph is generated to act as the input node for level two on which the four process are repeated.
+A new graph $G_i=(V_i, E_i)$ is formed for the clustering at level $i > 1$ after the clustering at level $i-1$.
+Each cluster formed at a given level $i-1$ is denoted as a new node.
+Hence the number of clusters at level $i-1$ form the number of nodes at level $i$.
+The intra cluster edges form the edge set $E_i$ of the graph $G_i$.
+
+\end{itemize} % 1
+
+\subsection{Partitioning Phase}
+After coarsening the graph, we can apply any traditional graph partitioning approaches to partitiong it. In our
+case, we are randomly select nodes and put them in partitions. While partitioning we also make sure that the
+size of does not differ by large quantity i.e they should be of nearly equal size.
+%\end{itemize} % 1
+
+\subsection{Refinement Phase}
+When the required number of partitions are formed, it gives us an approximate partition of the graph. To refine the partitions, 
+we use a modified form of Kernighan-Lin algorithm\cite{22}.
+
+    The Kernighan-Lin algorithm starts with initial bipartitions of the graph and then iteratively optimizes it using a greedy approach. The Kernighan-Lin algorithm is divided into 
+two phases, in the first phase, we consider a pair of nodes, with each nodes from different partitions.
+Now, a gain value $\delta$$Q$ is calculated for all possible pairs, which is the change in the number of edges when the nodes in the pair are swapped.
+The pair with the maximum  gain is marked for swapping so that they are never marked again in this phase.
+This continues until all the pairs are marked.
+In the next phase, the algorithm examine the sequence of the swapped pairs and the sequence which results into the maximum gain is selected and the actual swapping takes place.
+This results into a locally optimized graph and the two phases are repeated till there is no further gain in the edges between the partitions.
+
+    We modify one of the phases of  Kernighan-Lin algorithm for our criterion.
+To be precise, we change the definition of $\delta$$Q$ for us.
+In the Kernighan-Lin algorithm, $\delta$$Q$ is the change in the edges cuts between the partitions when the nodes are swapped.
+Let $a$, $b$ be two nodes that lie in different partitions.
+Let $I_a$ be the internal cost for $a$, i.e the neigbors of $a$ that lies in the same partition as of $a$ and  $E_a$ be the external cost for the node $a$, i.e the number of neighbors of $a$ that lie in the separate partition.
+Similarily $E_b$ and $I_b$ are the external cost and internal cost for the node $b$.
+Now let $D_a$ = $E_a$ - $I_a$ and $D_b$ = $E_b$ - $I_b$ , i.e the difference of external and internal costs.
+If $a$ and $b$ are interchanged then the gain i.e $\delta$$Q$ will be $D_a + D_b -2C_{a,b}$.
+We make changes in the internal, external and $C_{a,b}$ to fit our purpose of decreasing the probability of the two hops from one node lying in the different partitions.
+We define internal cost $I_{ap}$ as the probability of an edge of node $E_{ap}$ lying in the same partition as of $a$ ,external cost as probability of an edge of node $a$ lying in a separate partition.
+Hence,\\
+\begin{displaymath}I_{a}=\frac{I_a}{Degree\ of\ node\ a }\\\end{displaymath},
+\\\begin{displaymath}E_{a}=\frac{E_a}{Degree\ of\ node\ a}\\\end{displaymath} and\\
+\begin{displaymath}2C_{a,b}=\frac{1}{Degree\ of\ node\ a}+\frac{1}{Degree\ of\ node\ b}\\\end{displaymath}
+
+    As the original Kernighan-Lin algorithm focuses on reducing the edges between the partitions, our modified one reduces the probability of one hop of the nodes to lie in the separate partition. 
+      The rest of the kernighan-Lin algorithm works in the original way.
+Experiments shows that it not only reduces the probability of one hop to lie outside the partitions but also decreases the edge cuts.
+The same approach could be applied to the linear heuristics from the feduccia and matthews
+algorithm\cite{23} which can make it more efficient in terms of time taken to refine the partitions.
+
+\section{Experiments and Results}
+
+\section{Conclusion}
+\bibliographystyle{aaai} \bibliography{xbib}
 \end{document}

### 2012-12-14 CMPT771/project_report/paper.tex
1 day to go!!

@@
 \begin{quote}
-Online social networking has  become a phenomenon in the last
-few years. The growing usage of social networks has put the service providers
-under a lot of pressure to efficiently process the users queries.
- In social networks like orkut, most of the queries access the immediate neighbors of the current node and
-neighbors of its neighbors.  Existing graph partitioning approaches focus on reducing the number
- of inter server links, neglecting the query
-patterns. Our algorithm focuses on partitioning the graph so as to
-minimize the number of servers to be accessed in order to
- process a query. Our results show that the proposed algorithm gives useful partitioning of social networks,
-thereby reducing inter server accesses.
+Online social networking has  become a phenomenon in the last few years. The
+growing usage of social networks has put the service providers under a lot of
+pressure to efficiently process the users queries.  In social networks like
+orkut, most of the queries access the immediate neighbors of the current node
+and neighbors of its neighbors.  Existing graph partitioning approaches focus
+on reducing the number of inter server links, neglecting the query patterns.
+Our algorithm focuses on partitioning the graph so as to minimize the number of
+servers to be accessed in order to process a query. Our results show that the
+proposed algorithm gives useful partitioning of social networks, thereby
+reducing inter server accesses.
 
@@
 \section{Introduction}
-    In the last few years, the online social networks(OSNs) have grown in huge proportion with respect to the number of users as well as its utility.
-The amount of usage of OSNs exceeds e-mail usage nowadays, which is remarkable.
- OSNs are the most popular sites on the Internet.
-A study\footnote{http://www.news.com.au/technology/story/0,28348,25170441-5014239,00.html} reported that OSNs account for 10\% of all the Internet time worldwide.
-The growth can be appreciated by the fact that there was 1382\% increase in the user base of twitter in one month[Feb-Mar 09] \footnote{http://blog.nielsen.com/nielsenwire/onlinemobile/twitters-tweet-smell-of-success}.
-There can be variations of OSNs in terms of usability e.g On orkut, the users interact with each other by exchanging scraps or photographs, while on SLASHDOT, the users interact via blogs.
-Facebook, LinkedIn are other examples of OSNs.
+    In the last few years, the online social networks(OSNs) have grown in huge
+proportion with respect to the number of users as well as its utility. The
+amount of usage of OSNs exceeds e-mail usage nowadays, which is remarkable.
+OSNs are the most popular sites on the Internet. A
+study\footnote{http://www.news.com.au/technology/story/0,28348,25170441-5014239,00.html}
+reported that OSNs account for 10\% of all the Internet time worldwide. The
+growth can be appreciated by the fact that there was 1382\% increase in the
+user base of twitter in one month[Feb-Mar 09]
+\footnote{http://blog.nielsen.com/nielsenwire/onlinemobile/twitters-tweet-smell-of-success}.
+There can be variations of OSNs in terms of usability e.g On orkut, the users
+interact with each other by exchanging scraps or photographs, while on
+SLASHDOT, the users interact via blogs. Facebook, LinkedIn are other examples
+of OSNs.
 
-    The phenomenal growth in the size of OSNs has given rise to scalability issues that can be solved by distributing online social network graph among different servers.
-However the basis of partitioning  is a major issue.
-That is, the graph partitioning could be based on structural, geo-political or traffic properties, which are discussed in detail in \cite{arxviv} for twitter and orkut.
-In this paper, we introduce a new partitioning criterion called \texttt{hop based partitioning} which takes into consideration the typical nature of social network queries.
+    The phenomenal growth in the size of OSNs has given rise to scalability
+issues that can be solved by distributing online social network graph among
+different servers. However the basis of partitioning  is a major issue. That
+is, the graph partitioning could be based on structural, geo-political or
+traffic properties, which are discussed in detail in \cite{arxviv} for twitter
+and orkut. In this paper, we introduce a new partitioning criterion called
+\texttt{hop based partitioning} which takes into consideration the typical
+nature of social network queries.
 
@@
 
-Social network queries typically visit their one hop and two hop neighbors.
-The small world phenomenon links any two users on a social network with a distance of six hops.
-Hence, while partitioning a social network, it would be beneficial to maintain the one-hop and two-hop neighbors of a frequented query node on the same server\footnote{through the rest of the paper the server and partition terms are used interchangeably} as itself.
-Further, it is expected that any graph partitioning algorithm attempts to reduce the inter server edges and equally distribute the data across servers.
+Social network queries typically visit their one hop and two hop neighbors. The
+small world phenomenon links any two users on a social network with a distance
+of six hops. Hence, while partitioning a social network, it would be beneficial
+to maintain the one-hop and two-hop neighbors of a frequented query node on the
+same server\footnote{through the rest of the paper the server and partition
+terms are used interchangeably} as itself. Further, it is expected that any
+graph partitioning algorithm attempts to reduce the inter server edges and
+equally distribute the data across servers.
 
 For example, \cite{flickr} analyzed large-scale traces of information
-dissemination in the Flickr social network. Flickr allows users to create two types of links: links to favorite
-photos (called favorites in Flickr) and links to other users (called contacts in Flickr). Users may \texttt{favorite-mark} a photo
- and share interesting photos with others.
-\cite{flickr} refers to users who include a photo in their favorite photos list as fans of that photo.
-It was observed that for less popular photos, 91\% of all fans are within two hops of the uploaders.
-Further, for even for top popular photos, 81\% of all fans are within two hops of uploaders.
-\cite{OHM} mentions that most information is found one-hop away from the query node.
-According to a study\cite{4}, the maximum distance between two persons is six hops.
-Hence, it is a reasonable assumption that a user interacts primarily within his 2-3 hop radius and at most his six hop radius.
+dissemination in the Flickr social network. Flickr allows users to create two
+types of links: links to favorite photos (called favorites in Flickr) and links
+to other users (called contacts in Flickr). Users may \texttt{favorite-mark} a
+photo and share interesting photos with others. \cite{flickr} refers to users
+who include a photo in their favorite photos list as fans of that photo. It was
+observed that for less popular photos, 91\% of all fans are within two hops of
+the uploaders. Further, for even for top popular photos, 81\% of all fans are
+within two hops of uploaders. \cite{OHM} mentions that most information is
+found one-hop away from the query node. According to a study\cite{4}, the
+maximum distance between two persons is six hops. Hence, it is a reasonable
+assumption that a user interacts primarily within his 2-3 hop radius and at
+most his six hop radius.
 
-    In this project, we propose a novel probability based objective function and  an algorithm to find 
-the clusters which optimizes that. Given a social network graph $G$, our algorithm partitions the graph while 
-meeting the following constraints:
+    In this project, we propose a novel probability based objective function
+and  an algorithm to find the clusters which optimizes that. Given a social
+network graph $G$, our algorithm partitions the graph while meeting the
+following constraints:
 \begin{itemize}
@@
 
-The paper is organized as follows. The related
-works is shown in section 2, the objective function in exaplained in section 3, We describe our algorithm  in
- section 4 and  experimental results are shown in section 5.
- Finally, we conclude the paper in section 6.
+The paper is organized as follows. The related works is shown in section 2, the
+objective function in exaplained in section 3, We describe our algorithm  in
+section 4 and  experimental results are shown in section 5.  Finally, we
+conclude the paper in section 6.
 
 \section{Related Works}
-  The study of graph partitioning algorithms based on different properties of social networks has been proposed by \cite{arxviv}.
-In the recent past, graph partitioning for dealing with scalability issues and handling large datasets has been proposed by \cite{5,6}.
+  The study of graph partitioning algorithms based on different properties of
+social networks has been proposed by \cite{arxviv}. In the recent past, graph
+partitioning for dealing with scalability issues and handling large datasets
+has been proposed by \cite{5,6}.
 
-    With the increased usage of social networks, the scientists have become more interested in analyzing the properties of social networks. 
-Sociologists showed that the average path distance between two Americans is 6 hops \cite{4} and \cite{8} shows
-the small world effect analysis. 
-Study of a online social network  finds the small world behavior and local clustering
- property  \cite{9}.
- The small world property for offline social networks has been discussed in \cite{9,10,11} 
-and it was verified for the online social network \cite{12,13}. Analysis of different structural and scale free properties has also been extensively analyzed \cite{12}.
+    With the increased usage of social networks, the scientists have become
+more interested in analyzing the properties of social networks.  Sociologists
+showed that the average path distance between two Americans is 6 hops \cite{4}
+and \cite{8} shows the small world effect analysis.  Study of a online social
+network  finds the small world behavior and local clustering property
+\cite{9}.  The small world property for offline social networks has been
+discussed in \cite{9,10,11} and it was verified for the online social network
+\cite{12,13}. Analysis of different structural and scale free properties has
+also been extensively analyzed \cite{12}.
 
-    The traditional graph partitioning algorithms\cite{16,17,18} partition the graph into roughly equal sized partition  and focusses on minimizing the edge cuts.
- Further, the sociological approaches like hierarchical clustering based partitioning and natural division 
-based partitioning approaches are popularity known\cite{19}. Another genre of graph partitioning 
-algorithms that focus on producing quality partitions in less time are multilevel algorithms, which 
-include METIS\cite{20} and its modification for power law graphs\cite{21}. There are other partitioning 
-algorithms like Kernighan-Lin\cite{22} and it's linear time heuristics\cite{23} can also be used as 
-refinement algorithms.
+    The traditional graph partitioning algorithms\cite{16,17,18} partition the
+graph into roughly equal sized partition  and focusses on minimizing the edge
+cuts.  Further, the sociological approaches like hierarchical clustering based
+partitioning and natural division based partitioning approaches are popularity
+known\cite{19}. Another genre of graph partitioning algorithms that focus on
+producing quality partitions in less time are multilevel algorithms, which
+include METIS\cite{20} and its modification for power law graphs\cite{21}.
+There are other partitioning algorithms like Kernighan-Lin\cite{22} and it's
+linear time heuristics\cite{23} can also be used as refinement algorithms.
 
-    In this paper, we  compare our results with METIS\cite{20}.
-METIS focuses on minimizing the cuts between the partitions as well as into roughly equal partitions.
-The METIS algorithm is divided into three phases termed as coarsening, partitioning and uncoarsening/refinement.
- The main motivation behind the algorithm is to generate an approximate graph of the original graph, partitioning it and then retracting the coarsening phase while doing the refinement between each iteration.
- Our algorithm also follows a hierarchical clustering approach where repeated merging of clusters is done
- for some iteration and then the graph randomly partitioned with imposed contraints of having nearly equal 
-partitions. Finally, the refinement algorithm
- is applied on the clusters thus obtained. We differ from METIS, in terms of our partitioning criterion, 
-our novel way of clustering and partitioning, and  the modification and the method of refinement applied 
-which is applied only once.
+    In this paper, we  compare our results with METIS\cite{20}. METIS focuses
+on minimizing the cuts between the partitions as well as into roughly equal
+partitions. The METIS algorithm is divided into three phases termed as
+coarsening, partitioning and uncoarsening/refinement.  The main motivation
+behind the algorithm is to generate an approximate graph of the original graph,
+partitioning it and then retracting the coarsening phase while doing the
+refinement between each iteration.  Our algorithm also follows a hierarchical
+clustering approach where repeated merging of clusters is done for some
+iteration and then the graph randomly partitioned with imposed contraints of
+having nearly equal partitions. Finally, the refinement algorithm is applied on
+the clusters thus obtained. We differ from METIS, in terms of our partitioning
+criterion, our novel way of clustering and partitioning, and  the modification
+and the method of refinement applied which is applied only once.
 
@@
 
-   To calculate our objective function, we assume that there is an equal chance of each user of an 
-OSN to come online. However, we can easily modify our objective function if we know the probability of each user
-coming online.
+   To calculate our objective function, we assume that there is an equal chance
+of each user of an OSN to come online. However, we can easily modify our
+objective function if we know the probability of each user coming online.
 
- Given some partitions, an evaluation criteria should be able to give a good index of 
-the quality of partitions in terms of fraction of one and two hops lying
-outside the partition and should also give importance to the past interaction between the users of OSNs.
+ Given some partitions, an evaluation criteria should be able to give a good
+index of the quality of partitions in terms of fraction of one and two hops
+lying outside the partition and should also give importance to the past
+interaction between the users of OSNs.
 
@@
 
-        This probability index gives us an overall probability of a two hop query from a node to lie in the
- different partition. In the following section we will propose an algorithm which optimizes this objective 
-function.
+        This probability index gives us an overall probability of a two hop
+query from a node to lie in the different partition. In the following section
+we will propose an algorithm which optimizes this objective function.
 
@@
 3. Repeat step 1-3 for the desired number of iterations.\\
+
 \textbf{Phase 2 - Partitioning}\\
- Randomly allocate nodes into the desired number of clusters. We can use any traditional\\
- approach towards graph partitioning, in our algorithm  we are using a randomized one\\
+ Randomly allocate nodes into the desired number of clusters. We can use any\\
+traditional approach towards graph partitioning, in our algorithm  we are using\\
+a randomized one\\
 
@@
 
-In this section, we introduce our graph partitioning algorithm.
-A social network  partitioning algorithm is expected to reduce inter server accesses requiring neighbors and the neighbors of neighbors of a query nodes to be present in the same server as itself.
- To reduce the partitioning overhead, it is important that inter server links be reduced while trying to maximize the number of nodes for which its one and two hop neighbors be on the same server as itself.
- Further, in order to equally distribute the workload across servers and optimally use the servers used, it is
- essential that the partitions size should be approximately equal. Our algorithm attempts to fulfil 
- the above requirements while attempting to maximize the number of nodes for which a majority of its 
-one-hop neighbors and with next priority its 2-hop neighbors are included within the same server.
- Our algorithm is divided into three phases, namely the graph coarsening phase, partitioning 
-phase and the refinement phase. The graph coarsening phase reduces the size of the original graph by some
-factor by clustering the nodes \emph{intelligently}.  A weight function is used 
-in order to decide which nodes should be clustered  during the coarsening phase. In the partitioning phase, we 
-randomly generate partitions of the coarsened graph. The refinement phase next 
-reduces the inter partition edges by making slight modifications to the partitioned graph. We next give 
-definitions of commonly used terms followed by a detailed description of each of the phases.
+In this section, we introduce our graph partitioning algorithm. A social
+network  partitioning algorithm is expected to reduce inter server accesses
+requiring neighbors and the neighbors of neighbors of a query nodes to be
+present in the same server as itself.  To reduce the partitioning overhead, it
+is important that inter server links be reduced while trying to maximize the
+number of nodes for which its one and two hop neighbors be on the same server
+as itself.  Further, in order to equally distribute the workload across servers
+and optimally use the servers used, it is essential that the partitions size
+should be approximately equal. Our algorithm attempts to fulfil the above
+requirements while attempting to maximize the number of nodes for which a
+majority of its one-hop neighbors and with next priority its 2-hop neighbors
+are included within the same server.  Our algorithm is divided into three
+phases, namely the graph coarsening phase, partitioning phase and the
+refinement phase. The graph coarsening phase reduces the size of the original
+graph by some factor by clustering the nodes \emph{intelligently}.  A weight
+function is used in order to decide which nodes should be clustered  during the
+coarsening phase. In the partitioning phase, we randomly generate partitions of
+the coarsened graph. The refinement phase next reduces the inter partition
+edges by making slight modifications to the partitioned graph. We next give
+definitions of commonly used terms followed by a detailed description of each
+of the phases.
 
 \subsection{Definitions}
-\textbf {\it{Social Network Graph :} }The users in the social network
-form the nodes of the graph while the relationships between
-the users form the edges of the social network graph.\\
-\textbf{\it{k-hop neighbor :}} A node is said to be a k-hop neighbor of a
-query node if there are  $K$-edges separating the node
-from the query node. A one hop neighbor can also be a two hop neighbor and so on in case of triangles present in the graph
+\textbf {\it{Social Network Graph :} }The users in the social network form the
+nodes of the graph while the relationships between the users form the edges of
+the social network graph.\\ \textbf{\it{k-hop neighbor :}} A node is said to be
+a k-hop neighbor of a query node if there are  $K$-edges separating the node
+from the query node. A one hop neighbor can also be a two hop neighbor and so
+on in case of triangles present in the graph
 
@@
 
-The aim of this phase is to produce an approximate partitioning of the original graph. Our approach is a heirarchical one. 
-We start with the original graph and then we try to reduce the graph size by \emph{intelligently} clustering
-nodes at each level. After clustering nodes, we form a new graph with clusters from the previous level treated
-as nodes in the current level. After repeating the steps for some iterations, we apply a traditional graph
-partitioning algorithm to \emph{latest} graph. And then we do the refining on the produced clusters to get
-the desired output. In the following sections, we will describe our \emph{intelligent} clustering approach at
-each level and then we will continue our discussion with the explanation of the refining algorithm.
+The aim of this phase is to produce an approximate partitioning of the original
+graph. Our approach is a heirarchical one.  We start with the original graph
+and then we try to reduce the graph size by \emph{intelligently} clustering
+nodes at each level. After clustering nodes, we form a new graph with clusters
+from the previous level treated as nodes in the current level. After repeating
+the steps for some iterations, we apply a traditional graph partitioning
+algorithm to \emph{latest} graph. And then we do the refining on the produced
+clusters to get the desired output. In the following sections, we will describe
+our \emph{intelligent} clustering approach at each level and then we will
+continue our discussion with the explanation of the refining algorithm.
 \begin{itemize} % 1
@@
 
-\item \textbf{Node Weight Calculation:} In this project, we adopt a probability based weight function. 
-Nodes of higher weight values are later picked as seeds during the process of clustering.
+\item \textbf{Node Weight Calculation:} In this project, we adopt a probability
+based weight function.  Nodes of higher weight values are later picked as seeds
+during the process of clustering.
 
@@
 \item \textbf{Probability based weight function:}  
-In section 3, we described the $P_{x_i}^{2-hop}$ as the probability of two hops from $x_i^{th}$ to lie in separate
-partition. To calculate the weight for each node $x_i$, we calculate $P_{x_i}^{2-hop}$ assuming the nodes and its 
-neighbors are in one cluster and rest of the nodes are in separate cluster. Higher the weight, more are the
-chances that even if we have the node and its neighbors in one cluster, still there is high chance of two hops
-from node $x_i$ lying in separate cluster. So we start by clustering $\emph{low weight}$ nodes and we continue
- in ascending order.
+In section 3, we described the $P_{x_i}^{2-hop}$ as the probability of two hops
+from $x_i^{th}$ to lie in separate partition. To calculate the weight for each
+node $x_i$, we calculate $P_{x_i}^{2-hop}$ assuming the nodes and its neighbors
+are in one cluster and rest of the nodes are in separate cluster. Higher the
+weight, more are the chances that even if we have the node and its neighbors in
+one cluster, still there is high chance of two hops from node $x_i$ lying in
+separate cluster. So we start by clustering $\emph{low weight}$ nodes and we
+continue in ascending order.
 
@@
 
-\item \textbf{Clustering Algorithm:} The algorithm behaves like a hierarchical clustering algorithm where at each level clusters are formed by clustering the clusters of the previous level.
-The clustering algorithm varies based on the weight picking function used as described below.
+\item \textbf{Clustering Algorithm:} The algorithm behaves like a hierarchical
+clustering algorithm where at each level clusters are formed by clustering the
+clusters of the previous level. The clustering algorithm varies based on the
+weight picking function used as described below.
 
-For a given level $i$, the nodes of the graph at level $i$ are processed in decreasing order of weight.
-Initially the input graph forms the input of the clustering algorithm for level one.
-For each level $i >1$, a new graph is generated which is fed as input to the clustering algorithm.
+For a given level $i$, the nodes of the graph at level $i$ are processed in
+decreasing order of weight. Initially the input graph forms the input of the
+clustering algorithm for level one. For each level $i >1$, a new graph is
+generated which is fed as input to the clustering algorithm.
 
@@
 
-$Cluster$ $Size$ = ($Number$ $of$ $Nodes$ / $Number$ $of$ $partitions$) $+$/$-$ $deviation$ $value$.\\
+$Cluster$ $Size$ = ($Number$ $of$ $Nodes$ / $Number$ $of$ $partitions$) $+$/$-$
+$deviation$ $value$.\\
 
-where $deviation$ $value$= $x\%$ of ($Number$ $of$ $Nodes$ / $Number$ $of$ $partitions$) for some user defined value of $x$.\\
+where $deviation$ $value$= $x\%$ of ($Number$ $of$ $Nodes$ / $Number$ $of$
+$partitions$) for some user defined value of $x$.\\
 
-The user defined value $x$ in the equation above is referred to as the "limit value" through the rest of the paper.
+The user defined value $x$ in the equation above is referred to as the "limit
+value" through the rest of the paper.
 
-The size of any cluster is required to be less than the predefined $Cluster$ $Size$.
-If no node can be selected as a seed node as on selecting it would be giving rise to a cluster that exceeds the $cluster$ $size$, then some heuristics applied to cluster the remaining nodes in a graph.
+The size of any cluster is required to be less than the predefined $Cluster$
+$Size$. If no node can be selected as a seed node as on selecting it would be
+giving rise to a cluster that exceeds the $cluster$ $size$, then some
+heuristics applied to cluster the remaining nodes in a graph.
 
-A cluster that cannot accommodate further nodes is marked and the new graph formed ignores the node representing this cluster for further clustering.
+A cluster that cannot accommodate further nodes is marked and the new graph
+formed ignores the node representing this cluster for further clustering.
 
-\item \textbf{Graph Formation from Clusters:} For the first level before the algorithm starts, each node of the input graph is represented as a cluster in itself.
-After the first level of clustering at level one, a new graph is generated to act as the input node for level two on which the four process are repeated.
-A new graph $G_i=(V_i, E_i)$ is formed for the clustering at level $i > 1$ after the clustering at level $i-1$.
-Each cluster formed at a given level $i-1$ is denoted as a new node.
-Hence the number of clusters at level $i-1$ form the number of nodes at level $i$.
-The intra cluster edges form the edge set $E_i$ of the graph $G_i$.
+\item \textbf{Graph Formation from Clusters:} For the first level before the
+algorithm starts, each node of the input graph is represented as a cluster in
+itself. After the first level of clustering at level one, a new graph is
+generated to act as the input node for level two on which the four process are
+repeated. A new graph $G_i=(V_i, E_i)$ is formed for the clustering at level $i
+> 1$ after the clustering at level $i-1$. Each cluster formed at a given level
+$i-1$ is denoted as a new node. Hence the number of clusters at level $i-1$
+form the number of nodes at level $i$. The intra cluster edges form the edge
+set $E_i$ of the graph $G_i$.
 
@@
 \subsection{Partitioning Phase}
-After coarsening the graph, we can apply any traditional graph partitioning approaches to partitiong it. In our
-case, we are randomly select nodes and put them in partitions. While partitioning we also make sure that the
-size of does not differ by large quantity i.e they should be of nearly equal size.
+After coarsening the graph, we can apply any traditional graph partitioning
+approaches to partitiong it. In our case, we are randomly select nodes and put
+them in partitions. While partitioning we also make sure that the size of does
+not differ by large quantity i.e they should be of nearly equal size.
 %\end{itemize} % 1
@@
 \subsection{Refinement Phase}
-When the required number of partitions are formed, it gives us an approximate partition of the graph. To refine the partitions, 
-we use a modified form of Kernighan-Lin algorithm\cite{22}.
+When the required number of partitions are formed, it gives us an approximate
+partition of the graph. To refine the partitions, we use a modified form of
+Kernighan-Lin algorithm\cite{22}.
 
-    The Kernighan-Lin algorithm starts with initial bipartitions of the graph and then iteratively optimizes it using a greedy approach. The Kernighan-Lin algorithm is divided into 
-two phases, in the first phase, we consider a pair of nodes, with each nodes from different partitions.
-Now, a gain value $\delta$$Q$ is calculated for all possible pairs, which is the change in the number of edges when the nodes in the pair are swapped.
-The pair with the maximum  gain is marked for swapping so that they are never marked again in this phase.
-This continues until all the pairs are marked.
-In the next phase, the algorithm examine the sequence of the swapped pairs and the sequence which results into the maximum gain is selected and the actual swapping takes place.
-This results into a locally optimized graph and the two phases are repeated till there is no further gain in the edges between the partitions.
+    The Kernighan-Lin algorithm starts with initial bipartitions of the graph
+and then iteratively optimizes it using a greedy approach. The Kernighan-Lin
+algorithm is divided into two phases, in the first phase, we consider a pair of
+nodes, with each nodes from different partitions. Now, a gain value $\delta$$Q$
+is calculated for all possible pairs, which is the change in the number of
+edges when the nodes in the pair are swapped. The pair with the maximum  gain
+is marked for swapping so that they are never marked again in this phase. This
+continues until all the pairs are marked. In the next phase, the algorithm
+examine the sequence of the swapped pairs and the sequence which results into
+the maximum gain is selected and the actual swapping takes place. This results
+into a locally optimized graph and the two phases are repeated till there is no
+further gain in the edges between the partitions.
 
     We modify one of the phases of  Kernighan-Lin algorithm for our criterion.
-To be precise, we change the definition of $\delta$$Q$ for us.
-In the Kernighan-Lin algorithm, $\delta$$Q$ is the change in the edges cuts between the partitions when the nodes are swapped.
-Let $a$, $b$ be two nodes that lie in different partitions.
-Let $I_a$ be the internal cost for $a$, i.e the neigbors of $a$ that lies in the same partition as of $a$ and  $E_a$ be the external cost for the node $a$, i.e the number of neighbors of $a$ that lie in the separate partition.
-Similarily $E_b$ and $I_b$ are the external cost and internal cost for the node $b$.
-Now let $D_a$ = $E_a$ - $I_a$ and $D_b$ = $E_b$ - $I_b$ , i.e the difference of external and internal costs.
-If $a$ and $b$ are interchanged then the gain i.e $\delta$$Q$ will be $D_a + D_b -2C_{a,b}$.
-We make changes in the internal, external and $C_{a,b}$ to fit our purpose of decreasing the probability of the two hops from one node lying in the different partitions.
-We define internal cost $I_{ap}$ as the probability of an edge of node $E_{ap}$ lying in the same partition as of $a$ ,external cost as probability of an edge of node $a$ lying in a separate partition.
-Hence,\\
+To be precise, we change the definition of $\delta$$Q$ for us. In the
+Kernighan-Lin algorithm, $\delta$$Q$ is the change in the edges cuts between
+the partitions when the nodes are swapped. Let $a$, $b$ be two nodes that lie
+in different partitions. Let $I_a$ be the internal cost for $a$, i.e the
+neigbors of $a$ that lies in the same partition as of $a$ and  $E_a$ be the
+external cost for the node $a$, i.e the number of neighbors of $a$ that lie in
+the separate partition. Similarily $E_b$ and $I_b$ are the external cost and
+internal cost for the node $b$. Now let $D_a$ = $E_a$ - $I_a$ and $D_b$ = $E_b$
+- $I_b$ , i.e the difference of external and internal costs. If $a$ and $b$ are
+interchanged then the gain i.e $\delta$$Q$ will be $D_a + D_b -2C_{a,b}$. We
+make changes in the internal, external and $C_{a,b}$ to fit our purpose of
+decreasing the probability of the two hops from one node lying in the different
+partitions. We define internal cost $I_{ap}$ as the probability of an edge of
+node $E_{ap}$ lying in the same partition as of $a$ ,external cost as
+probability of an edge of node $a$ lying in a separate partition. Hence,\\
 \begin{displaymath}I_{a}=\frac{I_a}{Degree\ of\ node\ a }\\\end{displaymath},
-\\\begin{displaymath}E_{a}=\frac{E_a}{Degree\ of\ node\ a}\\\end{displaymath} and\\
-\begin{displaymath}2C_{a,b}=\frac{1}{Degree\ of\ node\ a}+\frac{1}{Degree\ of\ node\ b}\\\end{displaymath}
+\\\begin{displaymath}E_{a}=\frac{E_a}{Degree\ of\ node\ a}\\\end{displaymath}
+and\\ \begin{displaymath}2C_{a,b}=\frac{1}{Degree\ of\ node\
+a}+\frac{1}{Degree\ of\ node\ b}\\\end{displaymath}
 
-    As the original Kernighan-Lin algorithm focuses on reducing the edges between the partitions, our modified one reduces the probability of one hop of the nodes to lie in the separate partition. 
-      The rest of the kernighan-Lin algorithm works in the original way.
-Experiments shows that it not only reduces the probability of one hop to lie outside the partitions but also decreases the edge cuts.
-The same approach could be applied to the linear heuristics from the feduccia and matthews
-algorithm\cite{23} which can make it more efficient in terms of time taken to refine the partitions.
+    As the original Kernighan-Lin algorithm focuses on reducing the edges
+between the partitions, our modified one reduces the probability of one hop of
+the nodes to lie in the separate partition.  The rest of the kernighan-Lin
+algorithm works in the original way. Experiments shows that it not only reduces
+the probability of one hop to lie outside the partitions but also decreases the
+edge cuts. The same approach could be applied to the linear heuristics from the
+feduccia and matthews algorithm\cite{23} which can make it more efficient in
+terms of time taken to refine the partitions.
 

### 2012-12-14 CMPT771/survey_report/paper.tex
1 day to go!!

@@
 \usepackage{courier}
+\usepackage{graphicx}
 \frenchspacing
@@
 \begin{itemize}
-\item \textbf{Expansion}:  
-\item \textbf{Internal density}:  
-\item \textbf{Cut Ratio}:  
-\item \textbf{Normalized Cut}:  
-\item \textbf{Maximum-ODF(Out Degree Fraction)}  
-\item \textbf{Average-ODF}:  
-\item \textbf{Flake-ODF}: 
+\item \textbf{conductance}: $f(S) = \frac{c_S}{c m_S+c_S}$ measures the fraction of edges that points outside the
+cluster
+\item \textbf{Expansion}: $f(S) = \frac{c_S}{n_S}$ measures the number of edges per node that point outside the
+cluster
+\item \textbf{Internal density}: $f(S) = 1 - \frac{m_S}{n_S(n-n_S)}$ is the fraction of all possible edges leaving
+the cluster
+\item \textbf{Cut Ratio}: f(S) = $\frac{c_S}{n_S(n-n_S)}$ is the fraction of all possible edges leaving the cluster 
+\item \textbf{Normalized Cut}:  f(S) = $\frac{c_S}{2m_S+c_S} + \frac{c_S}{2(m-m_S)+c_S}$
 \end{itemize}
@@
 \subsection{Single-criterion scores}
+As described earlier Single criterian scores are the scores which uses one of
+the criterias to give a quantitative measure for the "goodness" of the
+communities. One of the most widely used criteria is Modularity
+\cite{newmod2006} which measure the number of within community edges, relative
+to a null model of a random graph with the same degree distribution.
+
+Here are few more single criterian scores
 \begin{itemize}
-\item \textbf{Modularity}:
-\item \textbf{Modularity ratio}:
-\item \textbf{Volume}:
-\item \textbf{Edges cut}:
+\item \textbf{Modularity}: $\frac{1}{4m}(m_S-E(m_S))$, where $E(m_S)$ is the expected number of edges between
+the nodes in set $S$ in a random graph with the same degree distribution.
+\item \textbf{Modularity ratio}: $\frac{m_S}{E(m_S)}$ is alternative definition of the modularity, where we take 
+the ratio of the number of edges between the nodes of $S$ and the expected number of such edges under the
+null-model
+\item \textbf{Volume}: $\sum_{u\in S}d(u)$ is sum of degrees of nodes in $S$
+\item \textbf{Edges cut}:$c_S$ is the number of edges needed to be removed to disconnect nodes in $S$ from the 
+rest of the network
 \end{itemize}
 
-\section{Algorithms}
+\section{Modularity}
 
-\cite{test123}
+\begin{figure}
+%\begin{minipage}[htb]{0.48\linewidth}
+\includegraphics[scale=0.25]{pics/bipartition.eps}
+\caption{Bisection of graph such that the number of edges between the partitions are minimized. The dotted line
+is the proposed solution.}
+\label{fig:bipart}
+%\end{minipage}
+\end{figure}
+\section{Traditional Approaches} In the following sections we will discuss some
+of the traditional approaches towards finding communities and then we will move
+on to discuss some of the more recent ones as well.
+
+\section{Graph
+Partitioning} The problem of graph partitioning consists of dividing a graph
+into $k$ groups of predefined size, such that the edges lying between the $k$
+partitions are minimized.
+
+In the graph partitioning algorithms you need to specify the number of
+partitions needed. Suppose, we don't specify it, then a trivial solution will
+be to group all the vertices into one cluster such that number of edges lying
+outside the partition is reduced to zero.  Also, specifying the size of the
+partitions is important as without it, the size of partitions formed will be
+highly unbalanced. Though we can avoid this problem by choosing a different
+objective function.
+
+Graph partitioning is quite ubiquitous in parallel computing, circuit
+partitioning and layout, and in solving partial differential equations as well.
+Most of the variations are NP-hard. As a result there are lots of heuristics
+that does pretty good job. Most of the algorithms perform a bisection of the
+graph. However, partitioning into more than two clusters can be achieved by
+iterative bisectioning.
+
+The \emph{Kernighan-Lin algorithm} \cite{kerni70} is one of the earliest
+algorithms proposed and is still widely used as a last step of various graph
+partitioning algorithm. The algorithm was discovered when they were working on
+the problem of partitioning electronic circuits onto boards. The algorithm is
+basically an optimization of an objective function $Q$, which represents the
+difference between the number of edges inside the modules and the number of
+edges lying between them.  The input to the algorithm is the initial partition
+of the graph in two clusters of predefined size.  The initial partitions can be
+randomly generated or can be the output of another graph partitioning
+algorithm. Then, equal number of vertices are swapped between the two
+partitions, so that $Q$ has the maximal increase. To avoid trapping in the
+local maxima of $Q$, the algorithm includes some swaps that decrease the
+function $Q$. After a series of swaps with positive and negative gains, the
+partition with the largest value of Q is selected and used as starting point of
+a new series of iterations.  The algorithm is known to converge after some
+iterations and output depends highly on the initial partition.  So this graph
+is mostly used to \emph{refine} the output of some other graph partitioning
+algorithms. The run time complexity of it is $O(n^2 log\ n)$. But on the sparse
+graph, it reduces to $O(n^2)$.
+
+Another widely used technique is the \emph{spectral bisection method}
+\cite{barnes82}, which is based on the properties of the spectrum of the
+Laplacian matrix. Spectral clustering will be discussed in detail later.  Just
+a quick note here, that it is quite fast as it depends on computing the
+eigenvectors of the Laplacian whose computation is relatively fast. The method
+gives good partitions, which can be further improved by applying Kernighan-Lin
+algorithm.
+
+In general, the algorithms used for graph partitioning are not good for finding
+communities, as it is necessary to provide the number of partitions and the
+size of partitions as an input, which is quite unintutive in case of finding
+communities. Instead, we would except a community finding algorithm to give
+that as an output.
+
+\begin{figure}
+%\begin{minipage}[htb]{0.48\linewidth}
+\includegraphics[scale=0.30]{pics/heiro.eps}
+\caption{A dendogram, or hierarchical tree. Horizontal cut correspond to
+partitions of the graph in communites} \label{fig:heiro}
+%\end{minipage}
+\end{figure}
+\section{Hierarchical Clustering} In general, we don't know much about the
+community strcuture of a graph. It is highly unlikely to know the number of
+communities in a graph as well the size of it. In such cases, the graph
+partitioning approaches can only be used on a hit and trial basis, where you
+experiment with the input to the partitioning algorithm. On the other hand, the
+graph may have a hierarchical structure, i.e it may display several levels of
+grouping of the vertices, with small clusters included within larger clusters,
+and so on. So in such cases a hierarchical algorithm, i.e an algorithm which
+reveals hierarchical structure might be useful.
+
+In most of the hierarchical algorithms, the starting point is the definition of
+a similarity measure between vertices. After a measure is chosen, one computes
+the similarity for each pair of vertices.  This gives a new n*n matrix X, the
+similarity matrix. Focus of the hierarchical clustering techniques is to find
+groups of vertices with high similarity, and can be broadly classified into two
+categories:
+
+\subsection{Agglomerative Approach}
+In this bottom-up approach, one starts from the vertices as separate
+clusters(singletons) and ends up with the graph as a unique cluster. The
+clusters with high mutual similarity are merged, so we need to define a measure
+that estimates how similar clusters are, our of the similarity matrix $X$. In
+single linkage clustering, the similarity between two groups is the minimum
+element $x_{ij}$, with $i$ in one group and $j$ in the other. On the other
+hand, in complete linkage clustering maximum element $x_{ij}$ is used. while in
+average linkage clustering we need to compute the average of $x_ij$.
+\ref{fig:heiro} gives a good illustration of a hierarchical algorithm. The red
+line in the figure denotes a stopping conditions which can be imposed based on
+the number of vertices in partitions or other objective function.
+
+The main advantage of using a hierarchical clustering is that it does not
+require a preliminary knowledge on the number and size of the clusters. On the
+other hand, it does not provide any stopping criterian on its own. Also, the
+results depends on the similarity criteria chosen. It also produces an
+\emph{artifical} hierarchial tree, even if a graph doesn't have any hierarchy
+in it. Also it is often the case that a node with only one neighbor forms a
+separate cluster, which is not good. However, the main disadvantage is it does
+not scale well.  Computing the similarity matrix is most expensive with best
+case of $O(n^2)$ in case of single linkage, while $O(n^2 log\ n)$ in other.
+
+\begin{figure}
+\includegraphics[scale=0.30]{pics/edgeb.eps}
+\caption{Edge betweenness is higest for edges connecting communities. As you
+can see in this figure the edge connecting the two communities will have much
+higher edge betweeness than the rest of the edges as all the shortest paths
+connecting vertices of the two communities run through it.} \label{fig:edge}
+\end{figure}
+
+\subsection{Divisive Approach}
+This is just the opposite of Agglomerative approach. Here we start with the
+original graph and consider them in one cluster and then iteratively remove
+edges based on some dissimilarity measure. The intution is, if we remove the
+edges that lie between the communities first, then we have our clusters.
+
+\subsubsection{Edge betweeness and Community structure}
+The most popular divisive hierarchical algorithm is proposed by Girvan and
+Newman\cite{new2002,new2004}. These were the seminal work which ignited new
+interest in the area of community detection and also opened this topic to
+physicists. In these papers, Girvan and Newman focussed on the concept of
+betweeness, which quantifies the participation of edges to a process. They
+considered edge betweeness, which is the number of shortest paths between all
+vertex pairs that run along the edge. The inutition behind this is the edges
+which lies between communities will have high edge betweeness value. So
+removing them will results into separation of the communities. This forms the
+basis of their algorithm. Here are the steps:-
+\begin{enumerate}
+\item Calculate the betweenness for all edges in the network
+\item Remove the edge with the highest betweenness
+\item Recalculate betweennesses for all edges affected by the removal
+\item Repeat from step 2 until no edges remain
+\end{enumerate}
+The naive algorithm doesn't scale very well. It is known that the worst case
+time complexity of this version is $O(n^3)$ on a sparse graph. Lot of heuristics have been
+proposed to scale this algorithm, one of the first was given by Wilkins et al \cite{wilkin2004}.
+
+\section{Partitional clustering}
+If we are given a set of data points then we can apply Partitional clustering
+to find clusters.  We are also give $k$ as the number of clusters we need to
+find. The points are embedded in a metric space. Each vertex is represented as
+a point and the distance between them is the measure of similarity between the
+vertices. The main aim of such kind of algorithms is to separate the points in
+$k$ clusters such as to optimize a given cost function based on distance
+between points. Some of the most popular cost functions are
+
+\begin{itemize}
+\item \textbf{Minimum k-clustering}: This represent the diameter of a cluster,
+which is the largest distance between two points of a cluster. The inution
+behind this is to keep the clusters compact so that the diameter is minimized.
+\item \textbf{k-clustering sum}:Same as above, but the diameter is replaced by
+the average distance between all pairs of points of a cluster.
+\item \textbf{k-center}:For each cluster $i$, we define a centroid and we
+compute the minimum $d_i$ of the distances of each point from the centroid of
+its cluster. The main aim is to choose clusters and centroids to minimize the
+largest value of $d_i$. 
+\item \textbf{k-media}:Same as k-center, except the distance  is replaced by
+average distance.
+\end{itemize}
+
+One of the most popular partitional algorithm is \emph{k-means clustering}
+\cite{kmeans67}. The cost function in this technique is the square error
+function given below. \\ \begin{equation} \sum_{i=1}^k \sum_{x_j \in S_i} ||x_j
+- c_i||^2 \end{equation} where $S_i$ indicates the set of points in i-th
+  cluster and $c_i$ is its centroid. This problem can be solved by using the
+algorithms proposed by Lloyd \cite{loyd82}. Given data points, we need to find
+$k$ clusters. The algorithm starts with choosing the $k$ centroids by some
+heuristics, most of them try to maximize the distance between them.
+In the first iteration, each vertex is assigned to the nrearest centroid. 
+Then the centroids for each cluster is estimated with the help points in the
+clusters.  Then the process of assigning the vertex to nearest centroid and
+restimating centroids is repeated till convergence. The final output depends
+heavily on the choice of the intial $k$ centroids. This is the main reason why
+people normally run this algorithms few times with different initialization and
+pick the clusters with minimum square function. The main reason why this
+heuristic is so popular is due to its quick convergence property which makes
+the analysis of large data sets fairly quick.
+
+Another technique which is quite famous and is related to k-means clustering is
+\emph{fuzzy k-means clustering} \cite{bez84}. This algorithm is widely used in pattern  recognition. 
+The main motivation behind building this algorithm is that a point may belong to two or more clusters
+at the same time. The cost function is
+
+\begin{equation}
+J_m = \sum_{i=1}^n \sum_{j=1}^k u_{ij}^m ||x_i-c_j||^2
+\end{equation}
+
+where $u_ij$ is the membership matrix, which measures the degree of membership
+of point $i$ (with position $x_i$) in cluster $j$, $m$ is a real number greater
+than 1 and $c_j$ is the center of cluster $j$
+ \begin{equation} c_j =
+\frac{\sum_{i=1}^n u_{ij}^m x_i}{\sum_{i=1}^n u_{ij}^m} 
+\end{equation}
+
+The matrix $u_{ij}$ is normalized such that the sum of the memberships of every
+point in all clusters yields 1. The membership $u_ij$ is related to the distance of point $i$ from the
+center of cluster $j$, as it is safe to assume that the larger this distance, the lowe $u_{ij}$. This
+can be expressed by the following relation
+
+\begin{equation}
+u_{ij}=\frac{1}{\sum_{l=1}^k(\frac{||x_i-c_j||}{||x_i-c_1||})^\frac{2}{m-l}}
+\end{equation}
+The cost function $J_m$ can be minimized by iterating Eq 3 and 4. Starting from
+initial guesses for $u_{ij}$ and using Eq. 3 to compute the centers, which are
+then used replaced in Eq 5, and so on. The process is stopped if the
+corresponding elements of the membership matrix in consecutive iterations
+differ from each other by less than some predefined value. This again depends
+on the initial choice of the matrix $u_{ij}$.
+
+The main disadvantage of partitional clustering is the same as that of graph
+partitioning algorithms i.e the number of clusters needs to be predefined and
+also one needs to find the embedding in the metric space if it not natural for
+some graphs.
+
+\section{Spectral Clustering} 
+Spectral clustering is another popular clustering technique which includes all
+methods and techniques that partition the set of nodes into cluster by using
+the eigenvectors of the similarity matrix of the nodes or other matrices which
+are derived from it. To calculate the eigenvectors the matrix needs to be
+symmetric and non-negative. Similarity matrix can be derived from the set of
+points in the metric space or from a graph as well. Spectral clustering mainly
+consists of a tranformation of the initial set of objects into a set of points
+in space, whose coordinates are elements of eigenvectors. The resulting points
+are then clustered using the standard clustering techniques, like k-means
+clustering. The main motivation behind this is due to the fact that the 
+change of representation induced by the eigenvectors make the cluster
+prpoerties of the initial data more evident. In this way, spectral clustering
+is able to separate data points into clusters which might not be very evident
+before applying this approach.
+
+\begin{figure}
+%\begin{minipage}[htb]{0.48\linewidth}
+\includegraphics[scale=0.25]{pics/spectral.eps}
+\caption{This figure shows a graph with 5 nodes and its Laplacian and the
+corresponding eigen values and eigen vectors} \label{fig:spectral}
+%\end{minipage}
+\end{figure}
+
+We will run through an example to show how to partition a graph using spectral
+clustering technique.  As given in \ref{fig:spectral}, we have a graph with
+five nodes. The first step is to calculate its Laplacian. After that we
+calculate the eigenvalues and corresponding eigenvectors for the laplacian. And
+then we sort the points according to the eigenvectors corresponding to second
+highest eigenvalue. As we can see it is able to separate the nodes in blue from
+nodes in red.
+
+In general there are three stages in spectral clustering:-
+\begin{itemize}
+\item Pre-processing
+\begin{itemize}
+\item Construct a matrix representation of the dataset
+\end{itemize}
+\item Decomposition
+\begin{itemize}
+\item Compute eigenvalues and eigenvectors of the matrix
+\item Map each point to a lower dimensional representation based on one or more
+eigenvectors
+\end{itemize}
+\item Grouping 
+\begin{itemize}
+\item Assign points to two or more clusters, based on the new representation
+\end{itemize}
+\end{itemize}
+
+\section{Methods based on statistical Inference}
+Statistical inference aims at deducing properties of data sets, starting from a
+set of observations and model hypotheses. In this section I will give a brief
+overview of how to use \textbf{generative models} to find clusters in a graph.
+Bayesian inference uses observations to estimate the probability that a given
+hypothesis is true.  It consists of two elements: first, the evidence,
+expressed by the information $D$ one has about the system and second, a
+statistical model with parameters $\{\theta\}$. In general, Bayesian inference
+start by writing the likelihood $P(D|\{\theta\})$ that the observed evidence is
+produced by the model for a given set of parameters $\{\theta\}$. The aim is to
+determine $\{\theta\}$ that maximizes the posterior distribution
+$P(\{\theta\}|D)$ of the parameters given the model and the evidence. By using Baye's theorem
+one has 
+\begin{equation}
+P(\{\theta\}|D) = \frac{1}{Z}P(D|\{\theta\}) P(\{\theta\})
+\end{equation}
+where $P(\{\theta\})$ is the prior distribution of the model parameters and
+
+\begin{equation}
+Z= \int P(D|\{\theta\}) P(\{\theta\}) d\theta
+\end{equation}
+The most challenging part is computing the above integral and also the choice
+of the prior distribution is not very obvious. Hence, most of the algorithms
+related to generative modesl differ from each other by the choice of the model
+and the way they addres these two issues.
+
+Bayesian inference is quite popular in the analysis and modelling of real
+graphs. Graph clustering can be considered a specific example of inference
+problem. The evidence is the graph structure and the hidden information is the
+classification of vertices into groups which one wants to infer. This idea
+forms the basis of several recent papers. Essentially, all of them maximizes
+the likelihood $P(D|\{\theta\})$ that the model is consistent with the observed
+graph structure.
+
 \bibliographystyle{aaai} \bibliography{bibfile}

### 2012-12-15 CMPT771/project_report/paper.tex
almost done

@@
 %
-\title{Probability Based Partitioning of Online Social Networks}
+\title{Hop Probability Based Partitioning of Online Social Networks}
 \author{Vivek Mahajan\\
@@
 pressure to efficiently process the users queries.  In social networks like
-orkut, most of the queries access the immediate neighbors of the current node
+facebook, most of the queries access the immediate neighbors of the current node
 and neighbors of its neighbors.  Existing graph partitioning approaches focus
@@
 \footnote{http://blog.nielsen.com/nielsenwire/onlinemobile/twitters-tweet-smell-of-success}.
-There can be variations of OSNs in terms of usability e.g On orkut, the users
-interact with each other by exchanging scraps or photographs, while on
-SLASHDOT, the users interact via blogs. Facebook, LinkedIn are other examples
+There can be variations of OSNs in terms of usability e.g On facebook, the users
+interact with each other by posting on the walls or sharing photographs, while on
+SLASHDOT, the users interact via blogs. LinkedIn and Google+ are other examples
 of OSNs.
@@
 issues that can be solved by distributing online social network graph among
-different servers. However the basis of partitioning  is a major issue. That
+different servers. However finding the basis of partitioning  is a major issue. That
 is, the graph partitioning could be based on structural, geo-political or
@@
 and orkut. In this paper, we introduce a new partitioning criterion called
-\texttt{hop based partitioning} which takes into consideration the typical
+\texttt{Hop probability based partitioning} which takes into consideration the typical
 nature of social network queries.
 
-\subsection{Hop Based Partitioning}
+\subsection{Hop probability Based Partitioning}
 
 Social network queries typically visit their one hop and two hop neighbors. The
-small world phenomenon links any two users on a social network with a distance
+small world phenomenon links any two users on a social network within a distance
 of six hops. Hence, while partitioning a social network, it would be beneficial
@@
 same server\footnote{through the rest of the paper the server and partition
-terms are used interchangeably} as itself. Further, it is expected that any
-graph partitioning algorithm attempts to reduce the inter server edges and
-equally distribute the data across servers.
+terms are used interchangeably} as itself.
 
@@
 
-    In this project, we propose a novel probability based objective function
+    In this project, we propose a novel hop probability based objective function
 and  an algorithm to find the clusters which optimizes that. Given a social
@@
     In this paper, we  compare our results with METIS\cite{20}. METIS focuses
-on minimizing the cuts between the partitions as well as into roughly equal
+on minimizing the cuts between the partitions as well as into roughly equal sized
 partitions. The METIS algorithm is divided into three phases termed as
@@
 clustering approach where repeated merging of clusters is done for some
-iteration and then the graph randomly partitioned with imposed contraints of
+iteration and then the graph is randomly partitioned with imposed contraints of
 having nearly equal partitions. Finally, the refinement algorithm is applied on
 the clusters thus obtained. We differ from METIS, in terms of our partitioning
-criterion, our novel way of clustering and partitioning, and  the modification
-and the method of refinement applied which is applied only once.
+criterion, our novel way of clustering, and  the modified form Kerinighan-Lin 
+algorithm which  refines the generated partitions.
 
-\section{Probability Based Objective Function}
+\section{Hop probability Based Objective Function}
 
@@
 of each user of an OSN to come online. However, we can easily modify our
-objective function if we know the probability of each user coming online.
+objective function if we know the probability of each user to come online.
 
- Given some partitions, an evaluation criteria should be able to give a good
-index of the quality of partitions in terms of fraction of one and two hops
+ Given partitions of a graph, our evaluation criteria gives a good
+quantitative measure of the quality of partitions in terms of fraction of one and two hops
 lying outside the partition and should also give importance to the past
-interaction between the users of OSNs.
-
-This index calculates the probability of one and two hops from a node to lie in different partition.
+interaction between the users of OSNs. Essentially, our objective function
+calculates the average probability of one and two hops from a node to lie in different
+partition.
 
 Formally, let $X$ = $\{x_1,x_2,x_3,\cdot\cdot\cdot,x_n\}$ be the nodes in an OSN, $x_i$ = $i^{th}\ node$, 
-$\{y_1, y_2, y_3,\cdot\cdot\cdot,y_j\}$ be the $j$ neighbors of $x_i$. Let $\{v_1, v_2, v_3, \cdot\cdot\cdot,
- v_j\}$ be the number of interactions between $x_i$ and its neighbors. And $V_i = \sum_{k\in j} v_k$.
- Let $Y^\prime$ be 
-the subset of neighbors of $x_i$ which lie in partition other than $x_i$'s.
-Let $Y^\prime = \{y_1^\prime, y_2^\prime, \cdot\cdot, y_k^\prime\}$.
-Let the interaction between $x_i$ and the users in this subset be  $V^\prime = \{v_1^\prime, v_2^\prime, \cdot\cdot, v_k^\prime\}$.
-$V_i^\prime = \sum_{p\in k} v_p^\prime$.
+$Y_i=\{y_1, y_2, y_3,\cdot\cdot\cdot,y_j\}$ be the $j$ neighbors of $x_i$. Let $V_i=\{v_1, v_2, v_3, \cdot\cdot\cdot,
+ v_j\}$ be the number of interactions between $x_i$ and its neighbors. And $|V_i| = \sum_{k\in j} v_k$.
+ Let $Y^\prime_i$ be 
+the subset of $Y_i$ which lie in partition other than $x_i$'s. Let $Y^\prime_i = \{y_1^\prime, y_2^\prime, \cdot\cdot, y_k^\prime\}$. 
+Let the interaction between $x_i$ and the users in this subset be  $V^\prime_i = \{v_1^\prime, v_2^\prime, \cdot\cdot, v_k^\prime\}$.
+$|V_i^\prime| = \sum_{p\in k} v_p^\prime$.
 
         Let $P_{x_i}^{1-hop}$ be the probability of one-hop neighbor of $x_i$ to lie in separate partition, and $P_{x_i}^{2-hop}$ be probability of two hops from $x_i^{th}$ node to lie in the separate partition.\\
-        Clearly, $P_{x_i}^{1-hop}= \frac{V_i^\prime}{V_i}$ \\
+        Clearly, $P_{x_i}^{1-hop}= \frac{|V_i^\prime|}{|V_i|}$ \\
         $P_{x_i}^{2-hop}=\frac{1}{Number\ of\ neighbors\ of\ x_i} * \displaystyle\sum_{k=1}^j P_{x_i y_k}$, \\
         where,\\
-        $P_{x_i y_k}$ = $\frac{v_k}{V_i}$* $P_{x_{k}}^{1-hop}$ =  Probability of two hop from  $x_{i}$ to lie  in a  partition other than  its own through $y_k$, if $x_i$ and $y_{k}$ are in the same partition.\\
+        $P_{x_i y_k}$ = $\frac{v_k}{|V_i|}$* $P_{x_{k}}^{1-hop}$ =  Probability of two hop from  $x_{i}$ to lie  in a  partition other than  its own through $y_k$, if $x_i$ and $y_{k}$ are in the same partition.\\
         $P_{x_i y_k}$ =  1 , if $x_i$ and $y_k$ are in the separate partition, as we assume if the one hop is on a different partition, it is almost
@@
         This probability index gives us an overall probability of a two hop
-query from a node to lie in the different partition. In the following section
-we will propose an algorithm which optimizes this objective function.
+query from a node to lie in the different partition. To the best of our
+knowledge this kind of evaluation of evaluation criteria has never been
+published before. This is quite novel in the sense it takes into account the
+user access patterns.  In the following section we will propose an algorithm
+which optimizes this objective function.
 
@@
 In this section, we introduce our graph partitioning algorithm. A social
-network  partitioning algorithm is expected to reduce inter server accesses
-requiring neighbors and the neighbors of neighbors of a query nodes to be
-present in the same server as itself.  To reduce the partitioning overhead, it
-is important that inter server links be reduced while trying to maximize the
-number of nodes for which its one and two hop neighbors be on the same server
-as itself.  Further, in order to equally distribute the workload across servers
-and optimally use the servers used, it is essential that the partitions size
-should be approximately equal. Our algorithm attempts to fulfil the above
-requirements while attempting to maximize the number of nodes for which a
-majority of its one-hop neighbors and with next priority its 2-hop neighbors
-are included within the same server.  Our algorithm is divided into three
+network  partitioning algorithm is expected to reduce inter server accesses,
+requiring friends and the friends of friends of a query node to be present in
+the same server as itself.   Also, in order to equally distribute the workload
+across servers , it is essential that the partitions size should be
+approximately equal.
+
+ Our algorithm attempts to fulfil the above requirements
+while attempting to maximize the number of nodes for which a majority of its
+one-hop neighbors and with next priority its 2-hop neighbors are included
+within the same server.  Our algorithm is divided into three
 phases, namely the graph coarsening phase, partitioning phase and the
@@
 coarsening phase. In the partitioning phase, we randomly generate partitions of
-the coarsened graph. The refinement phase next reduces the inter partition
-edges by making slight modifications to the partitioned graph. We next give
-definitions of commonly used terms followed by a detailed description of each
-of the phases.
+the coarsened graph. Due to time constraint we decided to randomly generate the
+partitions. Alternatively, we can use traditional graph partitioning methods.
+In the next phase, which we call refinement phase we reduce the inter partition
+edges by using a Kernighan-Lin algorithm. We next give definitions of commonly
+used terms followed by a detailed description of each of the phases.
 
@@
 the partitions when the nodes are swapped. Let $a$, $b$ be two nodes that lie
-in different partitions. Let $I_a$ be the internal cost for $a$, i.e the
-neigbors of $a$ that lies in the same partition as of $a$ and  $E_a$ be the
-external cost for the node $a$, i.e the number of neighbors of $a$ that lie in
-the separate partition. Similarily $E_b$ and $I_b$ are the external cost and
-internal cost for the node $b$. Now let $D_a$ = $E_a$ - $I_a$ and $D_b$ = $E_b$
+in different partitions. Let $I_a$ be the internal cost for $a$, i.e the total
+number of interactions between $a$ and its friends which lie in the same
+partition and  $E_a$ be the external cost for the node $a$, i.e the total
+number of interaction between $a$ and its friends which lie in the separate
+partition. Similarily $E_b$ and $I_b$ are the external cost and internal cost
+for the node $b$. Now let $D_a$ = $E_a$ - $I_a$ and $D_b$ = $E_b$
 - $I_b$ , i.e the difference of external and internal costs. If $a$ and $b$ are
@@
 node $E_{ap}$ lying in the same partition as of $a$ ,external cost as
-probability of an edge of node $a$ lying in a separate partition. Hence,\\
-\begin{displaymath}I_{a}=\frac{I_a}{Degree\ of\ node\ a }\\\end{displaymath},
-\\\begin{displaymath}E_{a}=\frac{E_a}{Degree\ of\ node\ a}\\\end{displaymath}
-and\\ \begin{displaymath}2C_{a,b}=\frac{1}{Degree\ of\ node\
-a}+\frac{1}{Degree\ of\ node\ b}\\\end{displaymath}
+probability of an edge of node $a$ lying in a separate partition. Let $T_a$ be
+$Total\ number\ of\ interactions\ done\ by\ user\ a$ and $T_b$ be $Total\
+number\ of\ interactions\ done\ by\ user\ b$. And $I_{ab}$ be the number of
+interactions between $a$ and $b$. Hence,\\
+\begin{displaymath}I_{ap}=\frac{I_a}{T_a}\\\end{displaymath},
+\\\begin{displaymath}E_{ap}=\frac{E_a}{T_a}\\\end{displaymath} and\\
+\begin{displaymath}2C_{ap,bp}=\frac{I_{ab}}{T_a}+\frac{I_{ab}}{T_b}\\\end{displaymath}
 
@@
 
+\begin{table*}[htb]
+\centering
+\caption{Datasets - statistics}
+\begin{tabular}{|c|c|c|c|}
+\hline
+ Datasets& Nodes & Edges \\
+\hline
+Facebook& 63731 & 68885  \\
+\hline
+SOC-Epinions & 75879 & 508837   \\
+\hline
+SLASHDOT& 77360 & 905468  \\
+\hline
+\end{tabular}
+
+\label{fig:twopartcuts}
+\end{table*}
+\begin{table*}[htb]
+\centering
+\caption{Two Partitions - Evaluation Indexes}
+\begin{tabular}{|c|c|c|c|c|c|c|}
+\hline
+ &\multicolumn{2}{c|}{Our Algorithm}&\multicolumn{2}{c|}{METIS}\\
+\hline
+%\cline{2-5}
+ DataSets&Hop Probability score &Edge Cuts&Hop Probability Score&Edge Cuts\\
+\hline
+Facebook& 0.225316 & 0.273173  & 0.21507 & 0.269943\\
+\hline
+SOC-Epinions & 0.135977 & 0.304285  & 0.162591 & 0.313931\\
+\hline
+SLASHDOT& 0.37827 & 0.496867  & 0.438342 & 0.762978\\
+\hline
+\end{tabular}
+\label{fig:twopartindex}
+\end{table*}
+
 \section{Experiments and Results}
+We have picked three popular and reasonaably sized datasets to run our experiments on
+\begin{itemize}
+\item \textbf{Facebook Social Network}: Facebook is a true social networking
+site on which users interact by writing on the wallposts and sharing lots of
+media content. The dataset was collected between December 29th, 2008 and
+January 3rd, 2009, and between January 20th 2009 and January 22th 2009 from the
+Facebook New Orleans network. The original dataset is time stamped, mentioning
+when were the links made.  For our purposes we used it as an undirected graph
+giving no consideration to time stamp attribute.  We also have list of all the
+wall posts for Facebook New Orleans network. The dataset was anonymized and we
+only have information about which user posted on whose wall at what time. No
+posts content is provided. But this information was good enough to find the
+interactions between the users which forms the weights of the edges between the
+users.  So we build a graph in which nodes are represented by users and there's
+and edge between two nodes if those two users are friends and weights are the
+average number of interactions between the two users i.e average number of
+times they posted on each other's wall + 1. 1 is added as there were cases when
+two users were friends but they didn't interact with each other. So we decided
+on having edge weight as 1 by default. As atleast they interacted once when
+they accepted the friends request. The dataset was crawled by \cite{26}.
 
-\section{Conclusion}
+\item \textbf{Epinions Social Network}: This is who-trust-whom online social
+network of a general consumer review site Epinions.com. Members of the site can
+decide whether to "trust" each other.  All the trust relationships interact and
+for the web of trust which is then combined with review rating to determine
+which reviews are shown to the user. We construct a graph in which the users
+are represented by nodes and there is an edge between a user if they "trust"
+each other. Since we didn't have any user query access pattern information for
+this graph, we gave each edge a weight of 1. The
+dataset\footnote{http://snap.stanford.edu/data/soc-Epinions1.html} was first
+used in \cite{epi2003}.
+
+\item \textbf{Slashdot}: Slashdot is a technology related news website known
+for its specific user community.  The website features user submitted and
+editor evaluated current primarily technology oriented news.  In 2002 Slashdot
+introduced the Slashdot Zoo feature which allows users to tag each other as
+neighbors or foes.  The network contains friend/foe links between the users of
+Slashdot. Again, the users are nodes and a link between them indicates whether
+they are friend or foe. And edge weight is set to 1 as we don't have any access
+pattern information. The
+dataset\footnote{http://snap.stanford.edu/data/soc-Slashdot0811.html} was first
+introduced in \cite{25}.
+
+\end{itemize}
+
+We compare the hop probability based partitioning approach with METIS. We are
+considering only bipartitions in our results. The input graph to the two
+algorithms are same except the facebook graph. For our algorithm, the input
+Facebook graph is with edge weights based on  interactions between users, while
+input Facebook graph for METIS is with all the edge weights set to 1. The main
+reason is that METIS doesn't take edge weights as input, so we can't provide it
+with that. We also have a reasonable contraint that the partitions should be of
+roughly equal size. We compare the two algorithms on edges cuts and hop
+probability based evaluation criteria. As we can see the periliminary results
+are quite close to METIS in terms of edge cuts as well the novel criteria. We
+think that some more experimentation with the weight function, partitioning
+algorithm(here we are doing random partitioning) and the refinement algorithm,
+might result in better results. 
+\section{Conclusion and Future Work}
+In this paper, our focus was to partition an Online Social Network graph such
+that the probability of query(one-hop/two-hop) generated from a user lying in
+different partition is reduced. We proposed a novel criteria for evaluating the
+partitions and presented a hierarchical algorithm optimizing that criteria. Our
+preliminary results looks promising. We plan to further improve it by 
+analyzing the graph formation at each level and also by experimenting with the
+weight function as well as the partitioning function. Refinement algorithm can
+be changed as well. As we don't know whether the modified Kernighan Lin
+algorithm will surely converge. I plan to work on this a little more and maybe
+submit it as a short paper in some conference/workshop.
 \bibliographystyle{aaai} \bibliography{xbib}

### 2012-12-15 CMPT771/survey_report/paper.tex
almost done

@@
 \begin{quote}
-Detecting clusters or communities in large real-world graphs such as large social or information networks is a problem of considerable interest. In general, it is the 
-organization of vertices in clusters, with many edges joining vertices of the same cluster and comparatively few edges joining vertices of different clusters. 
-In practice, one typically chooses an objective function that captures the intution as given by the definition above, and then one applies approximation algorithms
-or heuristics to extract sets of nodes which appears to be "good" with respect to that objective function.
+Detecting clusters or communities in large real-world graphs such as large
+social or information networks is a problem of considerable interest. In
+general, it is the organization of vertices in clusters, with many edges
+joining vertices of the same cluster and comparatively few edges joining
+vertices of different clusters.  In practice, one typically chooses an
+objective function that captures the intution as given by the definition above,
+and then one applies approximation algorithms or heuristics to extract sets of
+nodes which appears to be "good" with respect to that objective function.
 
-In this survey, we will start with the applications of clustering in various domains. We will describle various objective functions and will also discuss some of the
-most cited community finding papers.
+In this survey, we will start with the applications of clustering in various
+domains. We will describle various objective functions and will also discuss
+some of the most cited community finding papers.
 
@@
 \section{Introduction}
+The origin of graph theory dates back to Euler's solution of the puzzle of
+Konigsberg's bridges in 1736.  Since then a lot has been learned about graphs
+and their mathematical properties. It is arguably the most widely adopted
+representation which can be used to abstract lots of  real world problems. The
+problems in Biology, sociology and information networks can be formulated as
+graphs, and thus the graph analysis has become crucial to understand the
+features of these systems. Recently, with faster processors and plenty of
+computational resources, researchers can process and analyze huge amount of
+data, which wasn't possible in earlier times.
 
+Graphs representing real systems are more often not regular graphs. Some are
+random graph with higly homogeneous degree distribution. Real networks are not
+random graphs, as they display inhomogeneities, revealing a high level of order
+and organization, in which high degree vertices can co-exist with low degree
+vertices. Furthermore, the distribution of edges is not only globally, but also
+locally inhomogeneous, with high number of edges within some group of vertices,
+and low number of edges between these groups. This special feature of real
+networks is called community structure or clustering. Communities, also called
+clusters or modules, are groups of vertices which most likely share some common
+properties.
+
+\begin{figure}
+%\begin{minipage}[htb]{0.48\linewidth}
+\includegraphics[scale=0.40]{pics/intro.eps}
+\caption{A simple graph with three communities, enclosed by the dashed circles.}
+\label{fig:bipart}
+%\end{minipage}
+\end{figure}
+
+The communities exists in society, like families, working and friendship
+circles, towns, nations etc.  The ubiquity of Internet has also resulted into
+creation of virtual groups like online communities in social networks like
+Facebook or Linkedin. Communities also occur in many networked systems from
+biology, computer science, economics, politics etc.  In protein-protein
+interaction networks, communities are likely to group proteins having the same
+functional property in the cell. In WWW the may correspond to groups of pages
+dealing with same related topics, and so on.
+
+In WWW, clustering users who hvae similar interests and who are geografically
+near to each other may improve the performance of services provided on the
+World Wide Web, where each cluster of clients could be served by a dedicated
+mirror server. Clustering can be used  to generate reccomendations on
+e-commerce websites, e.g users with similar interests can be clustered and
+could be given appropriate reccomendation.  Clusters in large sparse graphs can
+be used to create data structures in order to efficiently store the graph data
+and to handle navigational queries, like path searches. It can also used in
+case of Ad hoc networks, as they normally don't maintain a central routing
+tables. Also, clustering can be used in case of caching of web pages as well.
+
+The goal of detecting communities in graphs is to identify the clusters and,
+more often, their hierarchical organization, by using the structure properties
+of the graph. The problem itself is very old and have lots of proposed
+algorithms in different disciplines.
+
+The first analysis of community structure was carried out by Weiss et al. in
+1957, who searched for group of workers in a government agency. Even earlier
+work was done in finding cluters of people in small political bodeis, based on
+the similarity of their voting pattern.
+
+Finding graph communities is also popular in computer science. In parallel
+computing, it is quite common to find out the best way to allocate tasks to
+processors so as to minimize the communications between them. This is also
+called graph partitioning problem which will be dicussed in the later sections.
+In a seminal paper appeared in 2002, Girvan and Newman proposed a new algorithm
+for finding the edges lying between communities and iteratively removing them
+to find the clusters \cite{new2002}. After this paper, the field was revived and people
+gave lots of sophisticated algorithms to tackle this problem.
+
+In this survery, we try to cover some of the most popular algorithms in this
+area. We start with giving some examples of communities in Real-world networks.
+In section 2, we will give definition of various objective functions used
+nowaday. We will continue the discussion with explaining one of the most widely
+used objective function in section 3. Followed by explanation of some
+traditional approaches in section 5, 6, 7 and 8. We will then give a brief
+overview of a more recent method in section 9 and then we will conclude in
+section 10.
+\subsection{Communities in Real-World Networks}
 \section{Objective Functions}
-If you have clusters of a given graph as an output of a clustering algorithm, we need to have some quantitative measure of how "good" they are. We call them objective
-function which the algorithms try to optimize. The intution behind most of the objective functions is that a good cluster has many edges internally and few edges 
-pointing outside. In this section, we look at the other objective functions that are based on this intution.
+If you have clusters of a given graph as an output of a clustering algorithm,
+we need to have some quantitative measure of how "good" they are. We call them
+objective function which the algorithms try to optimize. The intution behind
+most of the objective functions is that a good cluster has many edges
+internally and few edges pointing outside. In this section, we look at the
+other objective functions that are based on this intution.
 
-In general, the two criterias contributes heavily to the formulation of objective function. The first is the number of edges between the members of the clusters, and 
-second is the number of edges between the members of the cluster and remainder of the network. So, based on whether an objective function is using either one or both 
-criteria in its formulation, we can divide the objective functions into two categories\cite{leskovec2010}. The first group, that we refer to as Multi-criterion scores, combines both 
-criteria into a single objective function; while the second group of objective functions uses only  a single of the two criteria. We will give a quick overview of the
-various definitions of the objective functions which falls under the two categories.
+In general, the two criterias contributes heavily to the formulation of
+objective function. The first is the number of edges between the members of the
+clusters, and second is the number of edges between the members of the cluster
+and remainder of the network. So, based on whether an objective function is
+using either one or both criteria in its formulation, we can divide the
+objective functions into two categories\cite{leskovec2010}. The first group,
+that we refer to as Multi-criterion scores, combines both criteria into a
+single objective function; while the second group of objective functions uses
+only  a single of the two criteria. We will give a quick overview of the
+various definitions of the objective functions which falls under the two
+categories.
 
@@
 \item \textbf{conductance}: $f(S) = \frac{c_S}{c m_S+c_S}$ measures the fraction of edges that points outside the
-cluster
+cluster \cite{good04,shi2000}
 \item \textbf{Expansion}: $f(S) = \frac{c_S}{n_S}$ measures the number of edges per node that point outside the
-cluster
+cluster \cite{radi04}
 \item \textbf{Internal density}: $f(S) = 1 - \frac{m_S}{n_S(n-n_S)}$ is the fraction of all possible edges leaving
-the cluster
+the cluster \cite{radi04}
 \item \textbf{Cut Ratio}: f(S) = $\frac{c_S}{n_S(n-n_S)}$ is the fraction of all possible edges leaving the cluster 
-\item \textbf{Normalized Cut}:  f(S) = $\frac{c_S}{2m_S+c_S} + \frac{c_S}{2(m-m_S)+c_S}$
+\item \textbf{Normalized Cut}:  f(S) = $\frac{c_S}{2m_S+c_S} + \frac{c_S}{2(m-m_S)+c_S}$ \cite{shi2000}
 \end{itemize}

### 2012-12-16 CMPT771/project_report/paper.tex
saving the last rivision

@@
 The paper is organized as follows. The related works is shown in section 2, the
-objective function in exaplained in section 3, We describe our algorithm  in
+objective function in explained in section 3, We describe our algorithm  in
 section 4 and  experimental results are shown in section 5.  Finally, we
@@
 clustering approach where repeated merging of clusters is done for some
-iteration and then the graph is randomly partitioned with imposed contraints of
+iteration and then the graph is randomly partitioned with imposed constraints of
 having nearly equal partitions. Finally, the refinement algorithm is applied on
 the clusters thus obtained. We differ from METIS, in terms of our partitioning
-criterion, our novel way of clustering, and  the modified form Kerinighan-Lin 
-algorithm which  refines the generated partitions.
-
+criterion, our novel way of clustering, and  the modified form Kernighan-Lin 
+algorithm which  refines the generated partitions. In coarsening phase, our
+algorithm tries to collapse the nodes into clusters so as to minimize the
+evaluation criteria, while in METIS it doesn't take that into account. The refinement
+phase also focuses on minimizing the evaluation criteria. 
 \section{Hop probability Based Objective Function}
@@
 \textbf{Phase 2 - Refining}\\
- Refine the partitions using kenighan-Lin algorithm or the modified one.\\
+ Refine the partitions using Kernighan-Lin algorithm or the modified one.\\
 \hline
@@
 The aim of this phase is to produce an approximate partitioning of the original
-graph. Our approach is a heirarchical one.  We start with the original graph
+graph. Our approach is a hierarchical one.  We start with the original graph
 and then we try to reduce the graph size by \emph{intelligently} clustering
@@
 the two then $e_{ij}$ will be equal to $k$, for the first level $e_{ij}=number\ of\ interactions\ between\ v_i\ and\ v_j$.\footnote{number of interactions might the average number of times user $v_i$ and $v_j$ have interacted
-with eachother}
+with each other}
 
@@
 After coarsening the graph, we can apply any traditional graph partitioning
-approaches to partitiong it. In our case, we are randomly select nodes and put
+approaches to partition it. In our case, we are randomly select nodes and put
 them in partitions. While partitioning we also make sure that the size of does
@@
 
-    The Kernighan-Lin algorithm starts with initial bipartitions of the graph
+    The Kernighan-Lin algorithm starts with initial Bi-partitions of the graph
 and then iteratively optimizes it using a greedy approach. The Kernighan-Lin
@@
 number of interaction between $a$ and its friends which lie in the separate
-partition. Similarily $E_b$ and $I_b$ are the external cost and internal cost
+partition. $C_{a,b}$ is the average number of interactions
+between user $a$ and $b$. Similarily $E_b$ and $I_b$ are the external cost and internal cost
 for the node $b$. Now let $D_a$ = $E_a$ - $I_a$ and $D_b$ = $E_b$
@@
 interactions between $a$ and $b$. Hence,\\
-\begin{displaymath}I_{ap}=\frac{I_a}{T_a}\\\end{displaymath},
+\begin{displaymath}I_{ap}=\frac{I_a}{T_a},\\\end{displaymath}
 \\\begin{displaymath}E_{ap}=\frac{E_a}{T_a}\\\end{displaymath} and\\
@@
 between the partitions, our modified one reduces the probability of one hop of
-the nodes to lie in the separate partition.  The rest of the kernighan-Lin
+the nodes to lie in the separate partition.  The rest of the Kernighan-Lin
 algorithm works in the original way. Experiments shows that it not only reduces
@@
 edge cuts. The same approach could be applied to the linear heuristics from the
-feduccia and matthews algorithm\cite{23} which can make it more efficient in
+feduccia and Mathews algorithm\cite{23} which can make it more efficient in
 terms of time taken to refine the partitions.
@@
 %\cline{2-5}
- DataSets&Hop Probability score &Edge Cuts&Hop Probability Score&Edge Cuts\\
+ Datasets&Hop Probability score &Edge Cuts&Hop Probability Score&Edge Cuts\\
 \hline
-Facebook& 0.225316 & 0.273173  & 0.21507 & 0.269943\\
+Facebook& 0.042590 & 87642  & 0.031047 & 68315\\
 \hline
-SOC-Epinions & 0.135977 & 0.304285  & 0.162591 & 0.313931\\
+SOC-Epinions & 0.192315 & 62544  & 0.163753 & 55104\\
 \hline
-SLASHDOT& 0.37827 & 0.496867  & 0.438342 & 0.762978\\
+SLASHDOT& 0.398257 & 113356  & 0.438645 & 81465 \\
 \hline
@@
 \section{Experiments and Results}
-We have picked three popular and reasonaably sized datasets to run our experiments on
+We have picked three popular and reasonably sized datasets to run our experiments on
 \begin{itemize}
 \item \textbf{Facebook Social Network}: Facebook is a true social networking
-site on which users interact by writing on the wallposts and sharing lots of
+site on which users interact by writing on the wall-posts and sharing lots of
 media content. The dataset was collected between December 29th, 2008 and
@@
 two users were friends but they didn't interact with each other. So we decided
-on having edge weight as 1 by default. As atleast they interacted once when
-they accepted the friends request. The dataset was crawled by \cite{26}.
+on having edge weight as 1 by default. As at-least they interacted once when
+they accepted the friends request. The dataset was crawled by \cite{26}. We have chosen
+this dataset as it is one of the few Facebook datasets provided on the web, which also has
+the wall-posts information during that period. That gives us a close approximate of
+the users interaction amongst each other.
 
@@
 \end{itemize}
-
 We compare the hop probability based partitioning approach with METIS. We are
-considering only bipartitions in our results. The input graph to the two
+considering only Bi-partitions in our results. The input graph to the two
 algorithms are same except the facebook graph. For our algorithm, the input
 Facebook graph is with edge weights based on  interactions between users, while
-input Facebook graph for METIS is with all the edge weights set to 1. The main
-reason is that METIS doesn't take edge weights as input, so we can't provide it
-with that. We also have a reasonable contraint that the partitions should be of
+input Facebook graph for METIS is with all the edge weights set to 1.
+We also have a reasonable constraint that the partitions should be of
 roughly equal size. We compare the two algorithms on edges cuts and hop
-probability based evaluation criteria. As we can see the periliminary results
+probability based evaluation criteria.
+
+As we can see the preliminary results
 are quite close to METIS in terms of edge cuts as well the novel criteria. We
@@
 algorithm(here we are doing random partitioning) and the refinement algorithm,
-might result in better results. 
+might result in better results. Given the time span for the course project,
+we thought of running our algorithm on relatively small graphs. We wanted to check the performance
+in terms of edge cuts and the probability based evaluation criteron. We feel that
+since the first iteration of the algorithm gave us quite close results, we beleive further improvement
+over them will beat the more refined $METIS$ algorithm which took few years to develop.
+
 \section{Conclusion and Future Work}
@@
 be changed as well. As we don't know whether the modified Kernighan Lin
-algorithm will surely converge. I plan to work on this a little more and maybe
-submit it as a short paper in some conference/workshop.
+algorithm will surely converge. We also want to compare the algorithm 
+with variety of graph partitioning algorithm other than METIS and also design 
+an algorithm for generating $k$ partitions. I plan to work on this a little more and 
+ eventually submit it as a workshop paper in ICWSM'13.
 \bibliographystyle{aaai} \bibliography{xbib}

### 2012-12-16 CMPT771/survey_report/paper.tex
saving the last rivision

@@
 \maketitle
-\begin{abstract}
-\begin{quote}
+\section*{Abstract}
 Detecting clusters or communities in large real-world graphs such as large
@@
 vertices of different clusters.  In practice, one typically chooses an
-objective function that captures the intution as given by the definition above,
+objective function that captures the intuition as given by the definition above,
 and then one applies approximation algorithms or heuristics to extract sets of
@@
 In this survey, we will start with the applications of clustering in various
-domains. We will describle various objective functions and will also discuss
+domains. We will describe various objective functions and will also discuss
 some of the most cited community finding papers.
-
-\end{quote}
-\end{abstract}
 
@@
 Graphs representing real systems are more often not regular graphs. Some are
-random graph with higly homogeneous degree distribution. Real networks are not
+random graph with highly homogeneous degree distribution. Real networks are not
 random graphs, as they display inhomogeneities, revealing a high level of order
@@
 creation of virtual groups like online communities in social networks like
-Facebook or Linkedin. Communities also occur in many networked systems from
+Facebook or LinkedIn. Communities also occur in many networked systems from
 biology, computer science, economics, politics etc.  In protein-protein
@@
 
-In WWW, clustering users who hvae similar interests and who are geografically
+In WWW, clustering users who have similar interests and who are geographically
 near to each other may improve the performance of services provided on the
 World Wide Web, where each cluster of clients could be served by a dedicated
-mirror server. Clustering can be used  to generate reccomendations on
+mirror server. Clustering can be used  to generate recommendations on
 e-commerce websites, e.g users with similar interests can be clustered and
-could be given appropriate reccomendation.  Clusters in large sparse graphs can
+could be given appropriate recommendation.  Clusters in large sparse graphs can
 be used to create data structures in order to efficiently store the graph data
@@
 1957, who searched for group of workers in a government agency. Even earlier
-work was done in finding cluters of people in small political bodeis, based on
+work was done in finding clusters of people in small political bodies, based on
 the similarity of their voting pattern.
@@
 processors so as to minimize the communications between them. This is also
-called graph partitioning problem which will be dicussed in the later sections.
+called graph partitioning problem which will be discussed in the later sections.
 In a seminal paper appeared in 2002, Girvan and Newman proposed a new algorithm
@@
 
-In this survery, we try to cover some of the most popular algorithms in this
-area. We start with giving some examples of communities in Real-world networks.
+In this survey, we try to cover some of the most popular algorithms in this
+area. %We start with giving some examples of communities in Real-world networks.
 In section 2, we will give definition of various objective functions used
-nowaday. We will continue the discussion with explaining one of the most widely
+nowadays. We will continue the discussion with explaining one of the most widely
 used objective function in section 3. Followed by explanation of some
@@
 section 10.
-\subsection{Communities in Real-World Networks}
+%\subsection{Communities in Real-World Networks}
 \section{Objective Functions}
@@
 we need to have some quantitative measure of how "good" they are. We call them
-objective function which the algorithms try to optimize. The intution behind
+objective function which the algorithms try to optimize. The intuition behind
 most of the objective functions is that a good cluster has many edges
 internally and few edges pointing outside. In this section, we look at the
-other objective functions that are based on this intution.
+other objective functions that are based on this intuition.
 
@@
 \subsection{Single-criterion scores}
-As described earlier Single criterian scores are the scores which uses one of
+As described earlier Single criterion scores are the scores which uses one of
 the criterias to give a quantitative measure for the "goodness" of the
@@
 
-Here are few more single criterian scores
+Here are few more single criterion scores
 \begin{itemize}
@@
 \section{Modularity}
+One of the most popular objective function is Modularity \cite{new2004}. It is
+based on the intuition that a random graph is not expected to have a cluster
+structure. So one could tell the quality of the clusters by comparing the edge
+distribution inside the cluster with that of a random subgraph. This expected 
+expected edge density is given by a \emph{null model}. It is the original graph
+with some of its original structural properties but without any community structure.
+Formally, modularity is
+\begin{equation} Q = \frac{1}{2m}\sum_{ij}(A_{ij}-P_{ij})\delta(C_i,C_j),\end{equation}
+where $A$ is the adjacency matrix, $m$ is the total number of edges of the
+graph, and $P_{ij}$ is the expected number of edges between vertices $i$ and
+$j$ in the null model. The $\delta$ is 1 if vertices $i$ and $j$ are in the
+same community. The null model selection is subjective and lot of research has
+been done on it. One simple model could be that the graph keeps the same number
+of edges as the original graph, and that edges are placed with the same
+probability between any pair of vertices.
 
@@
 graph. However, partitioning into more than two clusters can be achieved by
-iterative bisectioning.
+iterative Bi-sectioning.
 
@@
 communities, as it is necessary to provide the number of partitions and the
-size of partitions as an input, which is quite unintutive in case of finding
+size of partitions as an input, which is quite unintuitive in case of finding
 communities. Instead, we would except a community finding algorithm to give
@@
 \caption{A dendogram, or hierarchical tree. Horizontal cut correspond to
-partitions of the graph in communites} \label{fig:heiro}
+partitions of the graph in communities} \label{fig:heiro}
 %\end{minipage}
@@
 \section{Hierarchical Clustering} In general, we don't know much about the
-community strcuture of a graph. It is highly unlikely to know the number of
+community structure of a graph. It is highly unlikely to know the number of
 communities in a graph as well the size of it. In such cases, the graph
@@
 average linkage clustering we need to compute the average of $x_ij$.
-\ref{fig:heiro} gives a good illustration of a hierarchical algorithm. The red
+Figure \ref{fig:heiro} gives a good illustration of a hierarchical algorithm. The red
 line in the figure denotes a stopping conditions which can be imposed based on
@@
 require a preliminary knowledge on the number and size of the clusters. On the
-other hand, it does not provide any stopping criterian on its own. Also, the
+other hand, it does not provide any stopping criterion on its own. Also, the
 results depends on the similarity criteria chosen. It also produces an
-\emph{artifical} hierarchial tree, even if a graph doesn't have any hierarchy
+\emph{artificial} hierarchical tree, even if a graph doesn't have any hierarchy
 in it. Also it is often the case that a node with only one neighbor forms a
@@
 \includegraphics[scale=0.30]{pics/edgeb.eps}
-\caption{Edge betweenness is higest for edges connecting communities. As you
+\caption{Edge betweenness is highest for edges connecting communities. As you
 can see in this figure the edge connecting the two communities will have much
-higher edge betweeness than the rest of the edges as all the shortest paths
+higher edge betweenness than the rest of the edges as all the shortest paths
 connecting vertices of the two communities run through it.} \label{fig:edge}
@@
 original graph and consider them in one cluster and then iteratively remove
-edges based on some dissimilarity measure. The intution is, if we remove the
+edges based on some dissimilarity measure. The intuition is, if we remove the
 edges that lie between the communities first, then we have our clusters.
 
-\subsubsection{Edge betweeness and Community structure}
+\subsubsection{Edge betweenness and Community structure}
 The most popular divisive hierarchical algorithm is proposed by Girvan and
@@
 physicists. In these papers, Girvan and Newman focussed on the concept of
-betweeness, which quantifies the participation of edges to a process. They
-considered edge betweeness, which is the number of shortest paths between all
-vertex pairs that run along the edge. The inutition behind this is the edges
-which lies between communities will have high edge betweeness value. So
+betweenness, which quantifies the participation of edges to a process. They
+considered edge betweenness, which is the number of shortest paths between all
+vertex pairs that run along the edge. The intuition behind this is the edges
+which lies between communities will have high edge betweenness value. So
 removing them will results into separation of the communities. This forms the
@@
 \item Remove the edge with the highest betweenness
-\item Recalculate betweennesses for all edges affected by the removal
+\item Recalculate betweenness for all edges affected by the removal
 \item Repeat from step 2 until no edges remain
@@
 \item \textbf{Minimum k-clustering}: This represent the diameter of a cluster,
-which is the largest distance between two points of a cluster. The inution
+which is the largest distance between two points of a cluster. The intuition
 behind this is to keep the clusters compact so that the diameter is minimized.
@@
 
-One of the most popular partitional algorithm is \emph{k-means clustering}
+One of the most popular Partitional algorithm is \emph{k-means clustering}
 \cite{kmeans67}. The cost function in this technique is the square error
@@
 heuristics, most of them try to maximize the distance between them.
-In the first iteration, each vertex is assigned to the nrearest centroid. 
+In the first iteration, each vertex is assigned to the nearest centroid. 
 Then the centroids for each cluster is estimated with the help points in the
 clusters.  Then the process of assigning the vertex to nearest centroid and
-restimating centroids is repeated till convergence. The final output depends
-heavily on the choice of the intial $k$ centroids. This is the main reason why
+reestimating centroids is repeated till convergence. The final output depends
+heavily on the choice of the initial $k$ centroids. This is the main reason why
 people normally run this algorithms few times with different initialization and
@@
 Another technique which is quite famous and is related to k-means clustering is
-\emph{fuzzy k-means clustering} \cite{bez84}. This algorithm is widely used in pattern  recognition. 
-The main motivation behind building this algorithm is that a point may belong to two or more clusters
-at the same time. The cost function is
+\emph{fuzzy k-means clustering} \cite{bez84}. This algorithm is widely used in
+pattern  recognition.  The main motivation behind building this algorithm is
+that a point may belong to two or more clusters at the same time. The cost
+function is
 
@@
 point in all clusters yields 1. The membership $u_ij$ is related to the distance of point $i$ from the
-center of cluster $j$, as it is safe to assume that the larger this distance, the lowe $u_{ij}$. This
+center of cluster $j$, as it is safe to assume that the larger this distance, the lower $u_{ij}$. This
 can be expressed by the following relation
@@
 
-The main disadvantage of partitional clustering is the same as that of graph
+The main disadvantage of Partitional clustering is the same as that of graph
 partitioning algorithms i.e the number of clusters needs to be predefined and
@@
 points in the metric space or from a graph as well. Spectral clustering mainly
-consists of a tranformation of the initial set of objects into a set of points
+consists of a transformation of the initial set of objects into a set of points
 in space, whose coordinates are elements of eigenvectors. The resulting points
@@
 change of representation induced by the eigenvectors make the cluster
-prpoerties of the initial data more evident. In this way, spectral clustering
+properties of the initial data more evident. In this way, spectral clustering
 is able to separate data points into clusters which might not be very evident
@@
 
-\begin{figure}
+\begin{figure*}
+\centering
 %\begin{minipage}[htb]{0.48\linewidth}
-\includegraphics[scale=0.25]{pics/spectral.eps}
+\includegraphics[scale=0.5]{pics/spectral.eps}
 \caption{This figure shows a graph with 5 nodes and its Laplacian and the
@@
 %\end{minipage}
-\end{figure}
+\end{figure*}
 
 We will run through an example to show how to partition a graph using spectral
-clustering technique.  As given in \ref{fig:spectral}, we have a graph with
+clustering technique.  As given in Figure \ref{fig:spectral}, we have a graph with
 five nodes. The first step is to calculate its Laplacian. After that we
-calculate the eigenvalues and corresponding eigenvectors for the laplacian. And
+calculate the eigenvalues and corresponding eigenvectors for the Laplacian. And
 then we sort the points according to the eigenvectors corresponding to second
@@
 of the prior distribution is not very obvious. Hence, most of the algorithms
-related to generative modesl differ from each other by the choice of the model
-and the way they addres these two issues.
+related to generative models differ from each other by the choice of the model
+and the way they address these two issues.
 
@@
 
+\section{Conclusion}
+In this survey paper, we give a brief overview of different kinds of community
+finding algorithms and also discussed the advantages and disadvantages
+associated with them. The scope of this survey limits us to the given
+algorithms. But there are lot of different kind of modularity optimization
+based algorithms and algorithms based on spin models given by physicists and so
+on.
+
+However, by looking at the advent of research in this field, one can say that
+it has grown in a rather chaotic way, without any precise direction. There is
+no theoretical framework that defines what clustering algorithms are proposed
+to do. Different people have different idea of what clusters should be. Most of
+them are consistent with each other, but as long as there is an disagreement,
+it is difficult to find the best algorithm to find the clusters. One can say,
+that the large number of research papers in this field might be due to above
+given fact.
+
+There has been lot of research in this area and it is still going strong. It is quite
+possible hthat the scientific community will eventually converge to a formal definition
+of community. Most of the algorithms these days produce similar results, but the main
+question is still open what's the best algorithm and have we found the best methods to
+evaluate them?
 \bibliographystyle{aaai} \bibliography{bibfile}
